### 2024-04-04 paper/paper.qmd
initial commit

---
title: "The Anatomy of a Hit: Statistically Learning from the Best"
subtitle: "Modeling the elements that make up top 10 artists' songs"
author: 
  - Luca Carnegie
thanks: "Code and data are available at: https://github.com/lcarnegie/valence-modeling."
date: April 18 2024
date-format: long
abstract: "We attempt to do some interesting things with models and music - stay tuned!"
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(spotifyr)

```

# Introduction

Throughout the history of the music industry, musicians successful or not have always pined for one thing - writing a hit and making it big. However, for large music firms, taking on musicians that have yet to generate considerable popularity is an large risk, since they are often gambling on an unknown je ne sais quoi, and often, any given artist is more likely to not produce a hit than produce a hit; this is easily shown through the presence of a few small "superstars" that consistently dominate the music market [@rosen1981]. This poses an interesting question for artists vying to become that superstar: what does make a song a "hit"?

A canonical way to measure how popular a song may be is through the 'valence' or perceived energy of a song. Though measuring valence has been done at a national and cross-artistic-work level [@dodds2010], there has never been any inquiry into song valence on an individual basis. Though difficult to exactly predict what individual song could be a hit, it is very possible to wean out insight from previously released music to statistically learn the most important elements of a great song.

Using data from Spotify's API, I construct a linear model of valence or "energy", using the discographies of the ten most popular artists [find a source] of every particular genre in every decade since 1950. I find that [there are interesting things to be found...]. Blah Blah Blah.

In the streaming era, competition for stardom is at an all-time high. Being cognizant of the most impactful elements could encourage artists to focus on what truly works to generate hits. This may allow them to carve their own paths outside of the controlling nature of a record contract [@burke1997]. These insights could be used to give more knowledge and freedom to artists to be able to strike out on their own and carve their own destinies in the world of music.

# Data {#sec-data}

# Model

## Model set-up

### Model justification

# Results

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point {#sec-second-point}

## Third discussion point {#sec-third-point}

## Weaknesses and next steps {#sec-weaknesses-next}

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
# #| eval: true
# #| echo: false
# #| message: false
# #| warning: false
# #| label: fig-ppcheckandposteriorvsprior
# #| layout-ncol: 2
# #| fig-cap: "Examining how the model fits, and is affected by, the data"
# #| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
# 
# pp_check(first_model) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(first_model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

\newpage

# References

### 2024-04-04 scripts/01-download_data.R

@@ -14,8 +14,21 @@ library(eFRED)
 
 # [...UPDATE THIS...]
 
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+#### Download Spotify Data for songs released since  ####
+
+beyonce <- get_artist_audio_features("beyonce")
+saveRDS(beyonce, "data/raw_data/beyonce.rds")
+
+beyonce <- 
+  readRDS(here::here("data/raw_data/beyonce.rds"))
+
+beyonce
+
+
+
+
+## Download FRED Data
+
 
 
 

### 2024-04-04 paper/paper.qmd
updated paper

@@
 title: "The Anatomy of a Hit: Statistically Learning from the Best"
-subtitle: "Modeling the elements that make up top 10 artists' songs"
+subtitle: "Modeling the most impactful elements of the most successful musicians."
 author: 
@@
 format: pdf
+toc: true
 number-sections: true
@@
 ```
+
+\newpage
 

### 2024-04-04 paper/paper.qmd
updated paper again

--- 2024-04-04 paper/paper.qmd
@@
 format: pdf
+toc: true
 number-sections: true
 bibliography: references.bib
+nocite: |
+  @*
 ---
@@
 
+\newpage
+
 # Introduction
@@
 
-A canonical way to measure how popular a song may be is through the 'valence' or perceived energy of a song.
-Though measuring valence has been done at a national and cross-artistic-work level [@dodds2010], there has never been any inquiry into song valence on an individual basis.
+One way to measure how popular a song is through the 'valence' or perceived energy of a song.[investigate the motivation here].
+Measuring valence has been done at a national and cross-artistic-work level [@dodds2010], but I investigate solely song valence on an individual basis.
 Though difficult to exactly predict what individual song could be a hit, it is very possible to wean out insight from previously released music to statistically learn the most important elements of a great song.
@@
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+@fig-stanareyouokay-2 is a hat plot. It shows... This suggests...
 
@@
 # References
+
+---
+
+---

### 2024-04-18 paper/paper.qmd
updated paper
Simulated dataset, scraped artist data from billboard.

@@
 title: "The Anatomy of a Hit: Statistically Learning from the Best"
-subtitle: "Modeling the most impactful elements of the most successful musicians."
+subtitle: "Modeling the most impactful elements of the most popular musicians."
 author: 
   - Luca Carnegie
-thanks: "Code and data are available at: https://github.com/lcarnegie/valence-modeling."
+thanks: "Code and data are available at: https://github.com/lcarnegie/popularity-modeling."
 date: April 18 2024
@@
 
-Throughout the history of the music industry, musicians successful or not have always pined for one thing - writing a hit and making it big.
-However, for large music firms, taking on musicians that have yet to generate considerable popularity is an large risk, since they are often gambling on an unknown je ne sais quoi, and often, any given artist is more likely to not produce a hit than produce a hit; this is easily shown through the presence of a few small "superstars" that consistently dominate the music market [@rosen1981].
+Throughout the history of the music industry, musicians successful or not have always pined for one thing - writing their first hit and making it big.
+However, for large music firms, taking on musicians that have yet to generate considerable popularity is a large risk, since they are often gambling on an unknown je ne sais quoi, and often, any given artist is more likely to not produce a hit than produce a hit; this is easily shown through the presence of a few small "superstars" that consistently dominate the music market [@rosen1981].
 This poses an interesting question for artists vying to become that superstar: what does make a song a "hit"?
@@
 # References
-
----
-
----

### 2024-04-18 paper/paper.tex
Simulated dataset, scraped artist data from billboard.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Anatomy of a Hit: Statistically Learning from the Best},
  pdfauthor={Luca Carnegie},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{The Anatomy of a Hit: Statistically Learning from the
Best\thanks{Code and data are available at:
https://github.com/lcarnegie/popularity-modeling.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Modeling the most impactful elements of the most popular
musicians.}
\author{Luca Carnegie}
\date{April 18, 2024}

\begin{document}
\maketitle
\begin{abstract}
We attempt to do some interesting things with models and music - stay
tuned!
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\newpage

\section{Introduction}\label{introduction}

Throughout the history of the music industry, musicians successful or
not have always pined for one thing - writing their first hit and making
it big. However, for large music firms, taking on musicians that have
yet to generate considerable popularity is a large risk, since they are
often gambling on an unknown je ne sais quoi, and often, any given
artist is more likely to not produce a hit than produce a hit; this is
easily shown through the presence of a few small ``superstars'' that
consistently dominate the music market (Rosen 1981). This poses an
interesting question for artists vying to become that superstar: what
does make a song a ``hit''?

One way to measure how popular a song is through the `valence' or
perceived energy of a song.{[}investigate the motivation here{]}.
Measuring valence has been done at a national and cross-artistic-work
level (Dodds and Danforth 2010), but I investigate solely song valence
on an individual basis. Though difficult to exactly predict what
individual song could be a hit, it is very possible to wean out insight
from previously released music to statistically learn the most important
elements of a great song.

Using data from Spotify's API, I construct a linear model of valence or
``energy'', using the discographies of the ten most popular artists
{[}find a source{]} of every particular genre in every decade since
1950. I find that {[}there are interesting things to be found\ldots{]}.
Blah Blah Blah.

In the streaming era, competition for stardom is at an all-time high.
Being cognizant of the most impactful elements could encourage artists
to focus on what truly works to generate hits. This may allow them to
carve their own paths outside of the controlling nature of a record
contract (Burke 1997). These insights could be used to give more
knowledge and freedom to artists to be able to strike out on their own
and carve their own destinies in the world of music.

\section{Data}\label{sec-data}

\section{Model}\label{model}

\subsection{Model set-up}\label{model-set-up}

\subsubsection{Model justification}\label{model-justification}

\section{Results}\label{results}

\section{Discussion}\label{discussion}

\subsection{First discussion point}\label{sec-first-point}

\subsection{Second discussion point}\label{sec-second-point}

\subsection{Third discussion point}\label{sec-third-point}

\subsection{Weaknesses and next steps}\label{sec-weaknesses-next}

Weaknesses and next steps should also be included.

\newpage

\appendix

\section*{Appendix}\label{appendix}
\addcontentsline{toc}{section}{Appendix}

\section{Additional data details}\label{additional-data-details}

\section{Model details}\label{sec-model-details}

\subsection{Posterior predictive
check}\label{posterior-predictive-check}

In \textbf{?@fig-ppcheckandposteriorvsprior-1} we implement a posterior
predictive check. This shows\ldots{}

In \textbf{?@fig-ppcheckandposteriorvsprior-2} we compare the posterior
with the prior. This shows\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# \#| eval: true}
\CommentTok{\# \#| echo: false}
\CommentTok{\# \#| message: false}
\CommentTok{\# \#| warning: false}
\CommentTok{\# \#| label: fig{-}ppcheckandposteriorvsprior}
\CommentTok{\# \#| layout{-}ncol: 2}
\CommentTok{\# \#| fig{-}cap: "Examining how the model fits, and is affected by, the data"}
\CommentTok{\# \#| fig{-}subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]}
\CommentTok{\# }
\CommentTok{\# pp\_check(first\_model) +}
\CommentTok{\#   theme\_classic() +}
\CommentTok{\#   theme(legend.position = "bottom")}
\CommentTok{\# }
\CommentTok{\# posterior\_vs\_prior(first\_model) +}
\CommentTok{\#   theme\_minimal() +}
\CommentTok{\#   scale\_color\_brewer(palette = "Set1") +}
\CommentTok{\#   theme(legend.position = "bottom") +}
\CommentTok{\#   coord\_flip()}
\end{Highlighting}
\end{Shaded}

\subsection{Diagnostics}\label{diagnostics}

\textbf{?@fig-stanareyouokay-1} is a trace plot. It shows\ldots{} This
suggests\ldots{}

\textbf{?@fig-stanareyouokay-2} is a hat plot. It shows\ldots{} This
suggests\ldots{}

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-alexander2023}
Alexander, Rohan. 2023. \emph{Telling Stories with Data}.
\url{https://tellingstorieswithdata.com/}.

\bibitem[\citeproctext]{ref-burke1997}
Burke, Andrew E. 1997. {``Small Firm Start-up by Composers in the
Recording Industry.''} \emph{Small Business Economics} 9 (6): 463--71.
\url{https://doi.org/10.1023/A:1007968604929}.

\bibitem[\citeproctext]{ref-dodds2010}
Dodds, Peter Sheridan, and Christopher M. Danforth. 2010. {``Measuring
the {Happiness} of {Large}-{Scale} {Written} {Expression}: {Songs},
{Blogs}, and {Presidents}.''} \emph{Journal of Happiness Studies} 11
(4): 441--56. \url{https://doi.org/10.1007/s10902-009-9150-9}.

\bibitem[\citeproctext]{ref-citeLahman}
Friendly, Michael, Chris Dalzell, Martin Monkman, and Dennis Murphy.
2020. \emph{Lahman: Sean {``Lahman''} Baseball Database}.
\url{https://CRAN.R-project.org/package=Lahman}.

\bibitem[\citeproctext]{ref-rstanarm}
Goodrich, Ben, Jonah Gabry, Imad Ali, and Sam Brilleman. 2022.
{``Rstanarm: {Bayesian} Applied Regression Modeling via {Stan}.''}
\url{https://mc-stan.org/rstanarm/}.

\bibitem[\citeproctext]{ref-kim2021}
Kim, Seon Tae, and Joo Hee Oh. 2021. {``Music Intelligence: {Granular}
Data and Prediction of Top Ten Hit Songs.''} \emph{Decision Support
Systems} 145 (June): 113535.
\url{https://doi.org/10.1016/j.dss.2021.113535}.

\bibitem[\citeproctext]{ref-citeR}
R Core Team. 2023. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-rosen1981}
Rosen, Sherwin. 1981. {``The {Economics} of {Superstars}.''} \emph{The
American Economic Review} 71 (5): 845--58.
\url{https://www.jstor.org/stable/1803469}.

\bibitem[\citeproctext]{ref-tidyverse}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\end{CSLReferences}

\end{document}

### 2024-04-18 scripts/00-simulate_data.R

@@ -1,19 +1,46 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates Spotify API dataset
+# Author: Luca Carnegie
+# Date: 18 April 2024
+# Contact: luca.carnegie@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(stringi)
+
+
+# Set seed for reproducibility
+set.seed(2024)
+
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+
+simulated_data <- tibble(
+  popularity = sample(0:100, 100, replace = TRUE), 
+  song = stri_rand_strings(100, 10), # Random strings of length 10 for song titles
+  artist = stri_rand_strings(100, 7), # Random strings of length 7 for artist names
+  energy = runif(100, 0, 1), # Uniform distribution between 0 and 1
+  valence = runif(100, 0, 1),
+  danceability = runif(100, 0, 1),
+  liveness = runif(100, 0, 1),
+  explicit = sample(0:1, 100, replace = TRUE), # Binary variable, assuming 0 is not explicit, 1 is explicit
+  instrumentalness = runif(100, 0, 1),
+  duration_ms = sample(150000:300000, 100, replace = TRUE), # Duration in milliseconds, example range
+  mode = sample(0:1, 100, replace = TRUE) # Binary variable, assuming 0 is minor, 1 is major
+)
+
+# View the first few rows of the simulated data
+head(simulated_data)
+
+
+## Simulate some graphs
+
+
+
+
 
 
 

### 2024-04-18 scripts/01-download_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from Spotify and FRED
+# Purpose: Downloads and saves data from the Spotify API
 # Author: Luca Carnegie
 # Date: 28 March 2023
 # Contact: luca.carnegie@mail.utoronto.ca
@@ -9,26 +9,30 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(rvest)
 library(spotifyr)
-library(eFRED)
 
-# [...UPDATE THIS...]
+## Scrape the Greatest of All Time Hot 100 Artists from Billboard and store them in a vector.
 
-#### Download Spotify Data for songs released since  ####
+raw_data <-
+  read_html(
+    "https://www.billboard.com/charts/greatest-hot-100-artists/"
+  )
 
-beyonce <- get_artist_audio_features("beyonce")
-saveRDS(beyonce, "data/raw_data/beyonce.rds")
+write(as.character(raw_data), "data/rawdata/greatestartists.html")
 
-beyonce <- 
-  readRDS(here::here("data/raw_data/beyonce.rds"))
+raw_data <- 
 
-beyonce
+names <- raw_data |> html_elements("h3") |> html_text() |> head(100)
+
+clean_names <- trimws(gsub("\\s+", " ", names))
+
+clean_names
 
 
+## For each artist in the vector, call the Spotify API for artist's top tracks and for audio features
 
-
-## Download FRED Data
-
+## Save each dataset as .csv or parquet? 
 
 
 

### 2024-04-20 scripts/00-simulate_data.R
downloaded all data!

@@ -10,6 +10,16 @@
 #### Workspace setup ####
 library(tidyverse)
 library(stringi)
+library(ggplot2)
+library(beepr)
+library(broom)
+library(broom.mixed)
+library(knitr)
+library(modelsummary)
+library(purrr)
+library(rstanarm)
+library(testthat)
+library(tidyverse)
 
 
 # Set seed for reproducibility
@@ -36,8 +46,84 @@ simulated_data <- tibble(
 head(simulated_data)
 
 
-## Simulate some graphs
+## Test simulated Data 
 
+# 1. Summary Statistics Check
+summary_stats <- simulated_data |> 
+  summarise(across(everything(), list(mean = mean, sd = sd, min = min, max = max)))
+print(summary_stats)
+
+# 2. Data Integrity Check (Verify that the data types and formats are correct)
+data_types <- sapply(simulated_data, class)
+print(data_types)
+
+# 3. Missing Values Test 
+missing_vals <- sum(is.na(simulated_data))
+print(missing_vals)
+
+# 4. Uniqueness Test
+duplicated_songs <- simulated_data$song |> 
+                           duplicated() |> 
+                           sum()
+
+duplicated_artists <- simulated_data$artist |> 
+                               duplicated() |> 
+                               sum()
+
+print(c(duplicated_songs, duplicated_artists))
+
+# 5. Range Validation Test
+range_checks <- sapply(simulated_data[c("energy", "valence", "danceability", "liveness", "instrumentalness")], 
+                       function(x) all(x >= 0 & x <= 1))
+print(range_checks)
+
+
+## Simulate some graphs (because the API kicked me out lol)
+
+## Liveness
+ggplot(simulated_data, aes(x = liveness)) +
+  geom_histogram(bins = 45, fill = "blue", color = "black") + # You can adjust the number of bins as needed
+  labs(title = "Histogram of Liveness",
+       x = "Liveness",
+       y = "Density") +
+  theme_minimal()
+
+## Popularity
+ggplot(simulated_data, aes(x = popularity)) +
+  geom_histogram(bins = 45, fill = "blue", color = "black") + # Adjust the number of bins as needed
+  labs(title = "Histogram of Popularity",
+       x = "Popularity",
+       y = "Density") +
+  theme_minimal()
+
+# Explicit or Not 
+
+ggplot(simulated_data, aes(x = explicit)) +
[…]

### 2024-04-20 scripts/01-download_data.R

@@ -10,35 +10,66 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rvest)
+library(xml2)
 library(spotifyr)
+library(usethis)
 
-## Scrape the Greatest of All Time Hot 100 Artists from Billboard and store them in a vector.
+#### Download Data ####
+
+# Scrape the Greatest of All Time Hot 100 Artists from Billboard and store them in a vector.
 
 raw_data <-
   read_html(
     "https://www.billboard.com/charts/greatest-hot-100-artists/"
   )
 
-write(as.character(raw_data), "data/rawdata/greatestartists.html")
+write_html(raw_data, 'data/raw_data/greatest100.html')
 
-raw_data <- 
+raw_data <- read_html('data/raw_data/greatest100.html')
 
-names <- raw_data |> html_elements("h3") |> html_text() |> head(100)
+names <- raw_data |> 
+          html_elements("h3") |> 
+          html_text() |> 
+          head(100)
 
-clean_names <- trimws(gsub("\\s+", " ", names))
+artists <- trimws(gsub("\\s+", " ", names))
 
-clean_names
+artists <- trimws(gsub("/", " and ", names))
 
 
-## For each artist in the vector, call the Spotify API for artist's top tracks and for audio features
+# Take the list of artists and get data for each one
 
-## Save each dataset as .csv or parquet? 
+# Initialize lists to store tibbles for each artist
+artist_audio_features_list <- list()
+artist_top_tracks_list <- list()
 
+# Loop through each artist and store their data
+for (artist_name in artists) {
+  # Get audio features
+  audio_features <- get_artist_audio_features(artist_name)
+  # Extract artist_id and use it to get top tracks
+  artist_id <- audio_features$artist_id[1] # Assuming 'artist_id' is correctly named and not null
+  
+  # Get top tracks for the US market
+  top_tracks <- get_artist_top_tracks(artist_id, market = "US")
+  
+  # Store data frames in respective lists
+  artist_audio_features_list[[artist_name]] <- audio_features
+  artist_top_tracks_list[[artist_name]] <- top_tracks
+  
+  Sys.sleep(sample(0:12)) #Stop for a random time to not get kicked out of API
+  print(paste("Got", artist_name, "'s data"))
+}
+
+
+# Merge lists of tibbles into single tibbles
[…]

### 2024-04-20 scripts/02-data_cleaning.R

@@ -1,44 +1,18 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw Spotify popularity/audio feature data and combines it 
+#          into a single dataset. 
+# Author: Luca Carnegie
+# Date: 19 April 2024
+# Contact: luca.carnegie@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Run 01-download_data.R to get the data
 
 #### Workspace setup ####
 library(tidyverse)
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+# Do the cleaning and merging based on two simulated datasets 
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+
 
 #### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-04-21 scripts/00-simulate_data.R
finished cleaning data.

@@ -28,25 +28,27 @@ set.seed(2024)
 
 #### Simulate data ####
 
+num_songs <- 100
+
 simulated_data <- tibble(
-  popularity = sample(0:100, 100, replace = TRUE), 
-  song = stri_rand_strings(100, 10), # Random strings of length 10 for song titles
-  artist = stri_rand_strings(100, 7), # Random strings of length 7 for artist names
-  energy = runif(100, 0, 1), # Uniform distribution between 0 and 1
-  valence = runif(100, 0, 1),
-  danceability = runif(100, 0, 1),
-  liveness = runif(100, 0, 1),
-  explicit = sample(0:1, 100, replace = TRUE), # Binary variable, assuming 0 is not explicit, 1 is explicit
-  instrumentalness = runif(100, 0, 1),
-  duration_ms = sample(150000:300000, 100, replace = TRUE), # Duration in milliseconds, example range
-  mode = sample(0:1, 100, replace = TRUE) # Binary variable, assuming 0 is minor, 1 is major
+  popularity = sample(0:100, num_songs, replace = TRUE), 
+  song = stri_rand_strings(num_songs, 10), # Random strings of length 10 for song titles
+  artist = stri_rand_strings(num_songs, 7), # Random strings of length 7 for artist names
+  energy = runif(num_songs, 0, 1), # Uniform distribution between 0 and 1
+  valence = runif(num_songs, 0, 1),
+  danceability = runif(num_songs, 0, 1),
+  liveness = runif(num_songs, 0, 1),
+  explicit = sample(0:1, num_songs, replace = TRUE), # Binary variable, assuming 0 is not explicit, 1 is explicit
+  instrumentalness = runif(num_songs, 0, 1),
+  duration_ms = sample(150000:300000, num_songs, replace = TRUE), # Duration in milliseconds, example range
+  mode = sample(0:1, num_songs, replace = TRUE) # Binary variable, assuming 0 is minor, 1 is major
 )
 
 # View the first few rows of the simulated data
 head(simulated_data)
 
 
-## Test simulated Data 
+## Test simulated Data using TestThat
 
 # 1. Summary Statistics Check
 summary_stats <- simulated_data |> 
@@ -105,6 +107,18 @@ ggplot(simulated_data, aes(x = explicit)) +
        y = "Count") +
   theme_minimal()
 
+
+# energy 
+
+# valence 
+
+# danceability 
[…]

### 2024-04-21 scripts/01-download_data.R

@@ -63,13 +63,16 @@ for (artist_name in artists) {
 
 
 # Merge lists of tibbles into single tibbles
-audio_features_tibble <- bind_rows(artist_audio_features_list, .id = "Artist_Name")
-top_tracks_tibble <- bind_rows(artist_top_tracks_list, .id = "Artist_Name")
+audio_features_spotify <- bind_rows(artist_audio_features_list, .id = "Artist_Name")
+top_tracks_tibble_spotify <- bind_rows(artist_top_tracks_list, .id = "Artist_Name")
+
+audio_features <- as_tibble(audio_features_spotify)
+top_tracks <- as_tibble(top_tracks_tibble_spotify)
 
 
 #### Save data ####
 
-write_csv(audio_features_tibble, "data/raw_data/audio_features.csv")
-write_csv(top_tracks_tibble, "data/raw_data/top_tracks.csv")
+write_csv(audio_features, "data/raw_data/audio_features.csv")
+write_csv(top_tracks, "data/raw_data/top_tracks.csv")
 
          

### 2024-04-21 scripts/02-data_cleaning.R

@@ -9,9 +9,55 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(janitor)
 
-# Do the cleaning and merging based on two simulated datasets 
 
+top_tracks <- read.csv("data/raw_data/top_tracks.csv")
+audio_features <- read.csv("data/raw_data/audio_features.csv")
+
+#Get columns from top_tracks
+
+#Clean the two CSVs
+cleaned_top_tracks <- top_tracks |> 
+                      select(Artist_Name, name, popularity) |>
+                      rename(song_name = name) |>
+                      clean_names() |>
+                      mutate(song_id = paste(artist_name, song_name, sep = " - "))
+                      
+                      
+                          
+cleaned_audio_features <- audio_features|>
+                          mutate(explicit_ = if_else(explicit, 1, 0)) |>
+                          select(artist_name, 
+                                 track_name,
+                                 energy,
+                                 valence,
+                                 danceability,
+                                 explicit_,
+                                 loudness,
+                                 duration_ms
+                                 ) |>
+                          rename(song_name = track_name) |>
+                          clean_names() |>
+                          mutate(song_id = paste(artist_name, song_name, sep = " - "))
+                          
+
+#Merge based on song name
+cleaned_ds <- cleaned_audio_features |>
+  # Perform an inner join on 'song_name' to filter only relevant songs from audio features
+  inner_join(cleaned_top_tracks, by = "song_id") 
+
+# There were duplicate songs, so drop ones with the lower popularity score 
+# After all, we are interested in what makes a song "more" popular. 
+
+# Remove duplicates by keeping the entry with the highest popularity score
+cleaned_ds <- cleaned_ds |>
+  group_by(song_id) |>
+  slice_max(order_by = popularity, with_ties = FALSE) |>
+  ungroup()
+
+# Check the top entries of the merged dataset to ensure correctness
+head(cleaned_ds)
 
 
 #### Save data ####

### 2024-04-21 scripts/03-test_data.R

@@ -10,6 +10,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
+## Write the tests using testthat. 
 
 #### Test data ####

### 2024-04-21 scripts/02-data_cleaning.R
tested data.

@@ -54,11 +54,30 @@ cleaned_ds <- cleaned_audio_features |>
 cleaned_ds <- cleaned_ds |>
   group_by(song_id) |>
   slice_max(order_by = popularity, with_ties = FALSE) |>
-  ungroup()
+  ungroup() |>
+  select(-artist_name.y, -song_name.y) |>
+  clean_names() |>
+  rename(
+    song_name = song_name_x, 
+    artist_name = artist_name_x
+  ) |>
+  select(
+    song_id, 
+    artist_name, 
+    song_name,
+    popularity, 
+    energy, 
+    valence, 
+    danceability, 
+    explicit, 
+    loudness, 
+    duration_ms
+  )
+  
 
 # Check the top entries of the merged dataset to ensure correctness
 head(cleaned_ds)
 
 
 #### Save data ####
-write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")
+write_csv(cleaned_ds, "data/analysis_data/analysis_data.csv")

### 2024-04-21 scripts/03-test_data.R

@@ -10,7 +10,49 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(testthat)
+library(readr)
 
-## Write the tests using testthat. 
 
-#### Test data ####
+# Load the dataset
+cleaned_ds <- read_csv("data/analysis_data/analysis_data.csv")
+
+# Begin defining tests
+test_that("Data loads correctly", {
+  expect_true(nrow(cleaned_ds) > 0)
+})
+
+test_that("Data columns are of correct type", {
+  expect_type(cleaned_ds$artist_name, "character")
+  expect_type(cleaned_ds$song_name, "character")
+  expect_type(cleaned_ds$energy, "double")
+  expect_type(cleaned_ds$valence, "double")
+  expect_type(cleaned_ds$danceability, "double")
+  expect_type(cleaned_ds$explicit, "double")
+  expect_type(cleaned_ds$loudness, "double")
+  expect_type(cleaned_ds$duration_ms, "double")
+  expect_type(cleaned_ds$song_id, "character")
+  expect_type(cleaned_ds$popularity, "double")
+})
+
+test_that("Numeric values are within expected ranges", {
+  expect_true(all(cleaned_ds$energy >= 0 & cleaned_ds$energy <= 1))
+  expect_true(all(cleaned_ds$valence >= 0 & cleaned_ds$valence <= 1))
+  expect_true(all(cleaned_ds$danceability >= 0 & cleaned_ds$danceability <= 1))
+  expect_true(all(cleaned_ds$explicit %in% c(0, 1)))
+  expect_true(all(cleaned_ds$popularity >= 0 & cleaned_ds$popularity <= 100))
+})
+
+test_that("Song IDs are unique", {
+  expect_true(length(unique(cleaned_ds$song_id)) == nrow(cleaned_ds))
+})
+
+test_that("There are no missing values in key columns", {
+  key_columns <- c("artist_name", "song_name", "song_id", "popularity")
+  for (col in key_columns) {
+    expect_true(all(!is.na(cleaned_ds[[col]])))
+  }
+})
+
+# You can run the tests with the following line
+test_dir("tests/")

### 2024-04-22 paper/paper.qmd
started to organize paper, ran some models to test, etc.

@@
 title: "The Anatomy of a Hit: Statistically Learning from the Best"
-subtitle: "Modeling the most impactful elements of the most popular musicians."
+subtitle: "An attempt to model"
 author: 
   - Luca Carnegie
-thanks: "Code and data are available at: https://github.com/lcarnegie/popularity-modeling."
+thanks: "Code and data are available at: https://github.com/lcarnegie/popularity-modeling.
+Special thanks go to Rajan Maghera, Oliver Cao, and Hannah Yu for the support and encouragement.
+You're all wonderful."
 date: April 18 2024
@@
 
+## Workflow {#sec-first-point}
+## Source
+## Measurement and Sampling 
+## Variables of Interest 
+### Energy
+
+```{r boilerplate-r-chunk}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: ALIGNMENT
+
+```
+
+### Valence
+
+```{r boilerplate-r-chunk}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: ALIGNMENT
+
+```
+
+### Danceability
+
+```{r boilerplate-r-chunk}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: ALIGNMENT
+
+```
+
+### Explicit Status
+
+```{r boilerplate-r-chunk}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: ALIGNMENT
+
+```
+
+### Loudness
+
+```{r boilerplate-r-chunk}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: ALIGNMENT
+
+```
+
+### Duration
+
+```{r boilerplate-r-chunk}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: ALIGNMENT
+
+```
+
 # Model
+
+Create a Correl matrix to motivate choosing of variables.
 
@@
 
+I guess put the results of the model?
+
 # Discussion
+
+What does everything mean?
 
@@
 ## Weaknesses and next steps {#sec-weaknesses-next}
+- Standard errors be huge. That means the regression is very imprecise. One way to
+resolve that would be to get a larger sample size. 
+- Sample size is kind of small - but that's the point? 
+Maybe constructing a fuller dataset with the entire discographies 
+of these people using Spotify's assistance would allow us to get more precise
+estimates of effects. 
+- People could repeat this experiment on a genre-based basis or using different 
+ratings of artists. Perhaps there are genre-specific relationships within other 
+variables in the spotify API that I haven't covered here.
 
-Weaknesses and next steps should also be included.
+\newpage
+
+# References
+
+::: {#refs}
+:::
 
@@
 
-# Additional data details
+# Model Testing
 
-# Model details {#sec-model-details}
+```{r boilerplate-r-chunk}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: ALIGNMENT
 
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-# #| eval: true
-# #| echo: false
-# #| message: false
-# #| warning: false
-# #| label: fig-ppcheckandposteriorvsprior
-# #| layout-ncol: 2
-# #| fig-cap: "Examining how the model fits, and is affected by, the data"
-# #| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-# 
-# pp_check(first_model) +
-#   theme_classic() +
-#   theme(legend.position = "bottom")
-# 
-# posterior_vs_prior(first_model) +
-#   theme_minimal() +
-#   scale_color_brewer(palette = "Set1") +
-#   theme(legend.position = "bottom") +
-#   coord_flip()
 ```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a hat plot. It shows... This suggests...
-
-\newpage
-
-# References

### 2024-04-22 scripts/02-data_cleaning.R

@@ -10,12 +10,11 @@
 #### Workspace setup ####
 library(tidyverse)
 library(janitor)
-
+library(arrow)
 
 top_tracks <- read.csv("data/raw_data/top_tracks.csv")
 audio_features <- read.csv("data/raw_data/audio_features.csv")
 
-#Get columns from top_tracks
 
 #Clean the two CSVs
 cleaned_top_tracks <- top_tracks |> 
@@ -47,22 +46,29 @@ cleaned_ds <- cleaned_audio_features |>
   # Perform an inner join on 'song_name' to filter only relevant songs from audio features
   inner_join(cleaned_top_tracks, by = "song_id") 
 
-# There were duplicate songs, so drop ones with the lower popularity score 
+# We lose a few songs from each artist in the inner join because of a disconnect 
+# between the top_songs dataset and the audio_features dataset - some songs on the
+# top 10 don't seem to have audio features that are query-able. 
+
+# There were also duplicates, so drop ones with the lower popularity score.  
 # After all, we are interested in what makes a song "more" popular. 
 
-# Remove duplicates by keeping the entry with the highest popularity score
 cleaned_ds <- cleaned_ds |>
   group_by(song_id) |>
-  slice_max(order_by = popularity, with_ties = FALSE) |>
+  slice_max(order_by = popularity, with_ties = FALSE) |> # Remove duplicates 
   ungroup() |>
-  select(-artist_name.y, -song_name.y) |>
+  select(-artist_name.y, -song_name.y) |> #drop artifacts from inner join
   clean_names() |>
+  mutate(
+    duration_ms = duration_ms/1000 #convert to seconds
+  ) |>
   rename(
     song_name = song_name_x, 
-    artist_name = artist_name_x
+    artist_name = artist_name_x,
+    duration_secs = duration_ms
   ) |>
   select(
-    song_id, 
+    song_id, #re-order the list to make it clean.
     artist_name, 
     song_name,
     popularity, 
@@ -71,8 +77,9 @@ cleaned_ds <- cleaned_ds |>
     danceability, 
     explicit, 
     loudness, 
-    duration_ms
+    duration_secs
   )
+  
   
 
 # Check the top entries of the merged dataset to ensure correctness
@@ -80,4 +87,10 @@ head(cleaned_ds)
 
 
 #### Save data ####
-write_csv(cleaned_ds, "data/analysis_data/analysis_data.csv")
+
[…]

### 2024-04-22 scripts/03-test_data.R

@@ -15,7 +15,7 @@ library(readr)
 
 
 # Load the dataset
-cleaned_ds <- read_csv("data/analysis_data/analysis_data.csv")
+cleaned_ds <- read_parquet("data/analysis_data/dataset.parquet")
 
 # Begin defining tests
 test_that("Data loads correctly", {
@@ -30,9 +30,9 @@ test_that("Data columns are of correct type", {
   expect_type(cleaned_ds$danceability, "double")
   expect_type(cleaned_ds$explicit, "double")
   expect_type(cleaned_ds$loudness, "double")
-  expect_type(cleaned_ds$duration_ms, "double")
+  expect_type(cleaned_ds$duration_secs, "double")
   expect_type(cleaned_ds$song_id, "character")
-  expect_type(cleaned_ds$popularity, "double")
+  expect_type(cleaned_ds$popularity, "integer")
 })
 
 test_that("Numeric values are within expected ranges", {

### 2024-04-22 scripts/04-model.R

@@ -1,32 +1,31 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models Spotify popularity scores as a function of various audio features
+# Author: Luca Carnegie
+# Date: April 21, 2024
+# Contact: luca.carnegie@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 01 -download_data.R, 02-data_cleaning.R
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(beepr)
+library(modelsummary)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/analysis_data/dataset.parquet")
+
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+
+model <-
+  lm(
+    popularity ~ energy + valence + danceability + explicit + loudness + duration_secs,
+    data = analysis_data
   )
 
+summary(sim_run_data_first_model)
 
 #### Save model ####
 saveRDS(

### 2024-04-22 paper/paper.qmd
wrote a good chunk of the paper. added mode again as a predictor.

@@
 title: "The Anatomy of a Hit: Statistically Learning from the Best"
-subtitle: "An attempt to model"
+subtitle: "Modeling Popularity on Spotify using API Audio Features"
 author: 
@@
 thanks: "Code and data are available at: https://github.com/lcarnegie/popularity-modeling.
-Special thanks go to Rajan Maghera, Oliver Cao, and Hannah Yu for the support and encouragement.
-You're all wonderful."
+Thank you, first, to Michaela Drouillard for your insights and help.
+Special hat tips go to Rajan Maghera and Oliver Cao for your support and encouragement.
+You're both awesome!"
 date: April 18 2024
@@
 
-```{r}
+```{r setup}
 #| include: false
@@
 library(tidyverse)
-library(spotifyr)
+library(ggplot2)
+library(arrow)
+library(kableExtra)
+library(knitr)
+
+spotify_data <- read_parquet("../data/analysis_data/dataset.parquet")
 
@@
 One way to measure how popular a song is through the 'valence' or perceived energy of a song.[investigate the motivation here].
-Measuring valence has been done at a national and cross-artistic-work level [@dodds2010], but I investigate solely song valence on an individual basis.
+Measuring the effects of specific features of art on populations has been done at a national and cross-artistic-work level [@dodds2010], but I investigate solely song valence on an individual basis.
 Though difficult to exactly predict what individual song could be a hit, it is very possible to wean out insight from previously released music to statistically learn the most important elements of a great song.
@@
 
-## Workflow {#sec-first-point}
-## Source
+This study was produced using a workflow based in R  [@citeR], with several additional packages facilitating data collection and analysis.
+Data was initially simulated using the tidyverse [@tidyverse], dplyr [@dplyr] and stringi [@stringi] packages.
+The names of Billboard Greatest Top 100 Artists [@greatestofhot100] were scraped using rvest [@rvest], saved with xml2 [@xml2], then fed into spotifyr [@spotifyr] to collect the audio and popularity data from the Spotify API.
+The tidyverse was further used to clean and merge the collected data into a single dataset and was saved in a parquet file using arrow [@arrow].
+Finally, the data was visualized and summarized with ggplot2 [@ggplot2] and modelsummary [@modelsummary].
+
 ## Measurement and Sampling
+
+To determine which artists' discographies to sample, I used Billboard Magazine's "Greatest of All Time Hot 100 Artists" list to isolate artists whose work has consistently performed the best on the Billboard Hot 100.
+First compiled in 1958, the Hot 100 is a weekly ranking that lists the most popular songs in the United States, tracking all music genres.
+The rankings are calculated from a combination of radio plays, song sales and streams/views on platforms like Spotify and YouTube [@aboutbillboard].
+Given the ranking's focus on music consumption, artists' whose music make it onto this list are expected to have discographies that garner a mass appeal and popularity, and thus have qualities that can be analyzed for patterns that could describe the elements of a "hit".
+
+To obtain quantitative measures of the elements in question, I turned to Spotify and it's API, which I accessed through the spotifyr R package.
+Through a free account, anyone can access fairly granular data on various audio features for most songs in their library, from tempo, to key, to generated metrics such as "danceability" and "speechiness", which are inferred algorithmically from the recorded audio features.
+
+A consolidated quantitative measure of the popularity of these artists' songs was also obtained through the Spotify API. According to Spotify's documentation,
+
+>The popularity is calculated by algorithm and is based, in the most part, on the total number of plays the track has had and how recent those plays are...
+[It is important to note that] the popularity value may lag actual popularity by a few days: the value is not updated in real time.
+
+Considering this, we can interpret that the popularity score provides a useful measure of a song's enduring popularity.
+Songs with high scores continue to be played frequently, even long after their release.
+Since Spotify is the leading music streaming service in the USA, with the largest customer base, their popularity metrics can reasonably align with the Billboard Top 100 rankings.
+This allows us to assess the relative lasting popularity of various songs using a relatively robust metric.
+
+## Data Source
+
+Other existing datasets were considered before conducting this analysis.
+In particular, Pham, Kyauk and Park's prior machine learning work using the Million Song dataset [@CITE THIS] created some contention for creating a new dataset.
+However, one key issue with the Million Song dataset is that without a robust measure for popularity, there is no way to infer meaningful relationships.
+Creating measures that are a good representation would require effort that has already been expended by Spotify.
+
+The appeals of using Spotify's metrics over the Million Song dataset are twofold: first, most songs on Spotify have a quantitative measure of popularity attached to them; second, this research is interested only in "the best" artists - if we regress the characteristics of the Million song dataset, there would be little meaningful information to be extracted, since parameters would be based on a large amount of data where most songs would not be popular, making them effectively useless in attempting to infer relationships between song features and popularity.
+
+The final dataset, after cleaning, is made up of 9 variables, with 773 observations (each song is an observation).
+To avoid issues with multicollinearity in audio features and confounding due to mismatching between sources, I deliberately prioritized using Spotify's algorithmically generated metric.
+To further scope my analysis and modelling, I narrowed my feature set provided by Spotify to energy, valence, danceability, presence of explicit lyrics, loudness, and song duration.
+
 ## Variables of Interest
-### Energy
 
-```{r boilerplate-r-chunk}
+There are 7 quantitative variables that are of interest in my constructed dataset.
+
+CORRECT THE GRAPHS' DETAILS.
+
+### Popularity
+
+```{r popularity-histogram}
 #| message: false
@@
 #| fig-cap: CAPTION
-#| fig-align: ALIGNMENT
+#| fig-align: center
+
+ggplot(spotify_data, aes(x = popularity)) +
+  geom_histogram(bins = 12, fill = "blue", color = "black") + 
+  labs(title = "Histogram of Popularity",
+       x = "Popularity",
+       y = "Density") +
+  theme_minimal() + 
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5))
 
@@
 ### Valence
+
+```{r scatter-valence}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: center
+
+# Scatter plot of Popularity vs. Energy
+ggplot(spotify_data, aes(x = valence, y = popularity)) +
+  geom_point(alpha = 0.6, color = "coral") +
+  ggtitle("Popularity vs. Valence") +
+  xlab("Valence") +
+  ylab("Popularity Score") +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5)
+  )
+
+```
 
@@
 
-```{r boilerplate-r-chunk}
+```{r scatter-danceability}
 #| message: false
@@
 #| fig-cap: CAPTION
-#| fig-align: ALIGNMENT
+#| fig-align: center
+
+# Scatter plot of Popularity vs. Energy
+ggplot(spotify_data, aes(x = danceability, y = popularity)) +
+  geom_point(alpha = 0.6, color = "coral") +
+  ggtitle("Popularity vs. Danceability") +
+  xlab("Danceability") +
+  ylab("Popularity Score") +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5)
+  )
+
+```
+
+```{r danceability-histogram}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: center
+
+ggplot(spotify_data, aes(x = danceability * 100)) +
+  geom_histogram(bins = 20, fill = "blue", color = "black") + 
+  labs(title = "Histogram of Danceability",
+       x = "Danceability (0-100%)",
+       y = "Density") +
+  theme_minimal() + 
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5))
+
+```
+
+### Mode
+
+```{r maj-min-barchart}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: center
+
+# Summarizing the data to count explicit vs. non-explicit songs
+song_counts <- spotify_data |>
+  group_by(mode) |>
+  summarise(count = n()) |>
+  mutate(explicit = if_else(mode == 1, "Major", "Minor"))
+
+# Plotting the counts of explicit and non-explicit songs
+ggplot(song_counts, aes(x = mode, y = count, fill = explicit)) +
+  geom_col() +  # geom_col is used for bar plots with predefined data summaries
+  ggtitle("Count of Major vs. Minor Songs") +
+  xlab("Song Mode") +
+  ylab("Number of Songs") +
+  scale_fill_manual(values = c("Major" = "skyblue", "Minor" = "salmon")) +
+  theme_minimal() +  # Adds a minimal theme for better aesthetics
+  theme(
+    axis.text.x = element_blank(),  # Hides the x-axis text
+    axis.ticks.x = element_blank(), 
+    plot.title.position = "plot",
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
+```{r density-plot}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: center
+
+# Comparing popularity distributions for different modes
+ggplot(spotify_data, aes(x = popularity, fill = as.factor(mode))) +
+  geom_density(alpha = 0.5) +
+  ggtitle("Popularity Distribution by Mode") +
+  xlab("Popularity Score") +
+  ylab("Density") +
+  scale_fill_manual(values = c("dodgerblue", "gold"), labels = c("Minor", "Major")) +
+  theme(
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
+```{r EVALUATE}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: center
+
+ggplot(spotify_data, aes(x = as.factor(mode), y = popularity, fill = as.factor(mode)))+
+  geom_boxplot() +
+  ggtitle("Effect of Song Mode on Popularity") +
+  xlab("Mode (0 = Minor Key, 1 = Major Key)") +
+  ylab("Popularity Score") +
+  scale_fill_manual(values = c("gray", "tomato")) + 
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5))
 
@@
 
-```{r boilerplate-r-chunk}
+```{r explicit-bar-chart}
 #| message: false
@@
 #| fig-cap: CAPTION
-#| fig-align: ALIGNMENT
+#| fig-align: center
+
+# Summarizing the data to count explicit vs. non-explicit songs
+song_counts <- spotify_data |>
+  group_by(explicit) |>
+  summarise(count = n()) |>
+  mutate(explicit = if_else(explicit == 1, "Explicit", "Not Explicit"))
+
+# Plotting the counts of explicit and non-explicit songs
+ggplot(song_counts, aes(x = explicit, y = count, fill = explicit)) +
+  geom_col() +  # geom_col is used for bar plots with predefined data summaries
+  ggtitle("Count of Explicit vs. Non-Explicit Songs") +
+  xlab("Song Type") +
+  ylab("Number of Songs") +
+  scale_fill_manual(values = c("Not Explicit" = "skyblue", "Explicit" = "salmon")) +
+  theme_minimal()  # Adds a minimal theme for better aesthetics
 
@@
 
-```{r boilerplate-r-chunk}
+```{r scatter-loudness}
 #| message: false
@@
 #| fig-cap: CAPTION
-#| fig-align: ALIGNMENT
+#| fig-align: center
+
+# Scatter plot of Popularity vs. Energy
+ggplot(spotify_data, aes(x = loudness, y = popularity)) +
+  geom_point(alpha = 0.6, color = "coral") +
+  ggtitle("Popularity vs. Loudness") +
+  xlab("Loudness") +
+  ylab("Popularity Score") +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5)
+  )
 
@@
 
-```{r boilerplate-r-chunk}
+```{r duration-histogram}
 #| message: false
@@
 #| fig-cap: CAPTION
-#| fig-align: ALIGNMENT
+#| fig-align: center
+
+ggplot(spotify_data, aes(x = duration_secs/60)) +
+  geom_histogram(bins = 20, fill = "blue", color = "black") + 
+  labs(title = "Histogram of Song Duration (Mins)",
+       x = "Duration (mins)",
+       y = "Density") +
+  theme_minimal() + 
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5))
+
+```
+
+```{r EVALUATE}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: center
+
+ggplot(spotify_data, aes(x = as.factor(explicit), y = popularity, fill = as.factor(explicit))) +
+  geom_boxplot() +
+  ggtitle("Effect of Explicit Content on Popularity") +
+  xlab("Explicit Content (0 = No, 1 = Yes)") +
+  ylab("Popularity Score") +
+  scale_fill_manual(values = c("gray", "tomato"))
 
@@
 # Model
-
-Create a Correl matrix to motivate choosing of variables.
 
@@
 resolve that would be to get a larger sample size. 
+- Billboard Data finishes in 2015: getting a more updated list from Billboard or somewhere else would be helpful. 
 - Sample size is kind of small - but that's the point? 
@@
 
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5))
+
 ```

### 2024-04-22 scripts/02-data_cleaning.R

@@ -29,11 +29,11 @@ cleaned_audio_features <- audio_features|>
                           mutate(explicit_ = if_else(explicit, 1, 0)) |>
                           select(artist_name, 
                                  track_name,
-                                 energy,
                                  valence,
                                  danceability,
                                  explicit_,
                                  loudness,
+                                 mode, 
                                  duration_ms
                                  ) |>
                           rename(song_name = track_name) |>
@@ -60,7 +60,8 @@ cleaned_ds <- cleaned_ds |>
   select(-artist_name.y, -song_name.y) |> #drop artifacts from inner join
   clean_names() |>
   mutate(
-    duration_ms = duration_ms/1000 #convert to seconds
+    duration_ms = duration_ms/1000,  #convert to seconds
+    explicit  = as.integer(explicit)
   ) |>
   rename(
     song_name = song_name_x, 
@@ -68,11 +69,10 @@ cleaned_ds <- cleaned_ds |>
     duration_secs = duration_ms
   ) |>
   select(
-    song_id, #re-order the list to make it clean.
     artist_name, 
     song_name,
     popularity, 
-    energy, 
+    mode, 
     valence, 
     danceability, 
     explicit, 

### 2024-04-22 scripts/03-test_data.R

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests variables for correctness proper format using testthat
+# Author: Luca Carnegie
+# Date: April 21 2023
+# Contact: luca.carnegie@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 01 -download_data.R and 02-data_cleaning.R
 
 
 #### Workspace setup ####
@@ -25,30 +24,25 @@ test_that("Data loads correctly", {
 test_that("Data columns are of correct type", {
   expect_type(cleaned_ds$artist_name, "character")
   expect_type(cleaned_ds$song_name, "character")
-  expect_type(cleaned_ds$energy, "double")
   expect_type(cleaned_ds$valence, "double")
   expect_type(cleaned_ds$danceability, "double")
-  expect_type(cleaned_ds$explicit, "double")
+  expect_type(cleaned_ds$explicit, "integer")
+  expect_type(cleaned_ds$mode, "integer")
   expect_type(cleaned_ds$loudness, "double")
   expect_type(cleaned_ds$duration_secs, "double")
-  expect_type(cleaned_ds$song_id, "character")
   expect_type(cleaned_ds$popularity, "integer")
 })
 
 test_that("Numeric values are within expected ranges", {
-  expect_true(all(cleaned_ds$energy >= 0 & cleaned_ds$energy <= 1))
   expect_true(all(cleaned_ds$valence >= 0 & cleaned_ds$valence <= 1))
   expect_true(all(cleaned_ds$danceability >= 0 & cleaned_ds$danceability <= 1))
   expect_true(all(cleaned_ds$explicit %in% c(0, 1)))
+  expect_true(all(cleaned_ds$mode %in% c(0, 1)))
   expect_true(all(cleaned_ds$popularity >= 0 & cleaned_ds$popularity <= 100))
 })
 
-test_that("Song IDs are unique", {
-  expect_true(length(unique(cleaned_ds$song_id)) == nrow(cleaned_ds))
-})
-
 test_that("There are no missing values in key columns", {
[…]

### 2024-04-22 scripts/04-model.R

@@ -9,19 +9,21 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(rstanarm)
-library(beepr)
 library(modelsummary)
 
 #### Read data ####
 analysis_data <- read_parquet("data/analysis_data/dataset.parquet")
+
+numeric_data <- analysis_data |> select(-song_id, -song_name, -artist_name)
+
+cor(numeric_data)
 
 
 ### Model data ####
 
 model <-
   lm(
-    popularity ~ energy + valence + danceability + explicit + loudness + duration_secs,
+    popularity ~  valence + danceability + explicit + mode + loudness + duration_secs,
     data = analysis_data
   )
 

### 2024-04-23 paper/paper.qmd
lots of work on paper: modeling especially.

--- 2024-04-22 paper/paper.qmd
@@
 ---
-title: "The Anatomy of a Hit: Statistically Learning from the Best"
-subtitle: "An attempt to model"
+title: "The Anatomy of a Hit"
+subtitle: "Modeling Popularity of Billboard Artists using Spotify Audio Features"
 author: 
@@
 thanks: "Code and data are available at: https://github.com/lcarnegie/popularity-modeling.
-Special thanks go to Rajan Maghera, Oliver Cao, and Hannah Yu for the support and encouragement.
-You're all wonderful."
-date: April 18 2024
+Thank you to Michaela Drouillard for your insights on Spotify and it's metrics.
+Special hat tip goes to Rajan Maghera and my amazing parents, Cristina and Shane for your constant support and encouragement.
+I could not have done this without you."
+date: April 24 2024
 date-format: long
-abstract: "We attempt to do some interesting things with models and music - stay tuned!"
+abstract: "What makes a hit pop song?
+This paper examines the audio features that characterize mainstream music's biggest hits.
+By analyzing data from the discographies of the highest-charting artists on the Billboard Hot 100, the research identifies key attributes associated with popular songs.
+Multivariate regression analysis reveals that higher levels of danceability, explicit lyrics, and loudness are positively related to popularity, while emotional positivity (valence) exhibits a negative relationship.
+Overall, this work quantifies some attributes underpinning iconic pop successes and can empower professional pop musicians to make more informed creative decisions with their work."
 format: pdf
@@
 
+# Load necessary libraries 
 library(tidyverse)
-library(spotifyr)
+library(ggplot2)
+library(arrow)
+library(kableExtra)
+library(knitr)
+library(broom)
+library(modelsummary)
+library(jtools)
+
+# Read in data for visualization
+spotify_data <- read_parquet("../data/analysis_data/dataset.parquet")
 
@@
 
-Throughout the history of the music industry, musicians successful or not have always pined for one thing - writing their first hit and making it big.
-However, for large music firms, taking on musicians that have yet to generate considerable popularity is a large risk, since they are often gambling on an unknown je ne sais quoi, and often, any given artist is more likely to not produce a hit than produce a hit; this is easily shown through the presence of a few small "superstars" that consistently dominate the music market [@rosen1981].
-This poses an interesting question for artists vying to become that superstar: what does make a song a "hit"?
+Throughout history, musicians have aspired to write hit songs and achieve mainstream success in the music industry.
+However, for major music companies, signing artists without a proven track record is a substantial risk, as most aspirants fail to produce hits.
+This has led to a concentrated market dominated by a small group of "superstars", such as Taylor Swift, who consistently top the charts [@rosen1981].
+Breaking into the music industry, then, presents a unique challenge for aspiring musicians' art: what elements contribute to a pop song's mass appeal?
 
-One way to measure how popular a song is through the 'valence' or perceived energy of a song.[investigate the motivation here].
-Measuring valence has been done at a national and cross-artistic-work level [@dodds2010], but I investigate solely song valence on an individual basis.
-Though difficult to exactly predict what individual song could be a hit, it is very possible to wean out insight from previously released music to statistically learn the most important elements of a great song.
+The field of Hit Song Science, which employs statistical methods to predict song popularity, has gained traction with the increasingly widespread availability of economical computing resources.
+Researchers [@dodds2010], students [@phametal2015], and likely record companies have attempted to construct models that attempt to explain song popularity, though the latter's efforts remain proprietary, leaving artists without access to this knowledge.
 
-Using data from Spotify's API, I construct a linear model of valence or "energy", using the discographies of the ten most popular artists [find a source] of every particular genre in every decade since 1950.
-I find that [there are interesting things to be found...].
-Blah Blah Blah.
+This study diverges from previous approaches by constructing a dataset solely from the discographies of the best-performing artists on the Billboard Hot 100, the preeminent metric of success in the United States music industry.
+By focusing exclusively on the most successful artists' music, the elements that define enduring hits in the US market can be more easily identified.
+The analysis employs multivariate regression to estimate a song's Spotify-generated popularity score, using several audio features derived from Spotify's API as predictors.
 
-In the streaming era, competition for stardom is at an all-time high.
-Being cognizant of the most impactful elements could encourage artists to focus on what truly works to generate hits.
-This may allow them to carve their own paths outside of the controlling nature of a record contract [@burke1997].
-These insights could be used to give more knowledge and freedom to artists to be able to strike out on their own and carve their own destinies in the world of music.
+The regression results reveal statistically significant positive relationships between danceability, explicit lyrics, loudness, and Spotify popularity score, as well as a significant negative relationship between song valence (general positivity) and popularity score, after controlling for other variables.
 
-# Data {#sec-data}
+In the current era of music streaming, competition for stardom is intense.
+Understanding the key elements driving song success could empower artists to craft hits more effectively and potentially achieve greater independence from record labels [@burke1997].
+This data-driven approach could provide a strategic advantage for artists in the pursuit of mainstream success.
 
-## Workflow {#sec-first-point}
-## Source
-## Measurement and Sampling 
-## Variables of Interest 
-### Energy
+The paper is structured as follows: [@sec-data] provides a detailed overview of the dataset and analyzes each predictor individually; [@sec-model] outlines the regression model; [@sec-results] presents and discusses the results; and [@sec-discussion] critically examines the findings and their implications.
 
-```{r boilerplate-r-chunk}
+# Data Analysis {#sec-data}
+
+The Billboard Hot 100 ranks the most popular U.S.
+songs weekly based on radio plays, sales, and streaming [@aboutbillboard].
+Artists with songs frequently on this chart should have mass appeal, motivating the analysis of elements that define the popularity of those "hits".
+Using Billboard also scopes this analysis to solely the US market
+
+Unlike other datasets like The Million Song dataset used by Pham et al., Spotify's web API provides ready-made quantitative audio features (tempo, key, danceability, etc.) as well as calculated popularity score based on recent and total play counts for artists' songs, facilitating analysis of popularity determinants.
+
+Since the popularity score is based on both recency of streams and total number of streams, continually high scores indicate frequent plays long after release.
+This means this score can be thought of measuring a song's enduring popularity.
+As the leading U.S.
+music streamer, Spotify's metrics can reasonably align with Billboard rankings, providing a relatively robust measure of a song's lasting popularity, which is crucial in attempting to infer the elements that make up a "hit".
+
+To analyze the audio feature determinants of popularity, I used R [@citeR], the tidyverse [@tidyverse] and related packages.
+Billboard's "Greatest of All Time Hot 100 Artists" [@greatestofhot100] list identified popular artists.
+The audio features of the Billboard artists' songs and popularity data were then acquired from Spotify's API via the spotifyr [@spotifyr] package.
+Spotify data was current as of 2024, while Billboard rankings were from 2015.
+The data was cleaned using the dplyr [@dplyr] and janitor [@janitor] packages and saved using arrow [@arrow].
+Variables were chosen based on expected impact on popularity, resulting in a dataset with 773 songs.
+
+For this analysis, the popularity score, valence (musical 'positivity'), danceability, mode (major or minor), presence of explicit lyrics, loudness, and song duration were made of interest.
+A sample of the cleaned dataset is shown in [@tbl-sample].
+This and other tables were created using the knitr [@knitr] packages.
+Visualizations and modelling summaries used ggplot2 [@ggplot2] and modelsummary [@modelsummary].
+
+```{r tbl-sample}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Popularity Scores and Audio Features
+#| fig-align: center
+
+#fix formatting before submitting
+
+spotify_data <- read_parquet("../data/analysis_data/dataset.parquet")
+
+sample_table <- spotify_data |>
+  rename(
+    `Artist Name` = artist_name,
+    `Song Name` = song_name,
+    Popularity = popularity,
+    Mode = mode,
+    Valence = valence, 
+    Danceability = danceability, 
+    `Explicit` = explicit, 
+    `Loudness (dB)` = loudness, 
+    `Duration (seconds)` = duration_secs
+  ) |>
+  head(5) |>
+  arrange(`Artist Name`)
+
+  kable(sample_table, format = "markdown") |>
+  kable_styling(full_width = F, font_size = 1) |>
+  column_spec(2, width = "1cm")  # Control the width of specific columns
+```
+
+The primarily categorical dataset was first analyzed variable-by-variable.
+Each variable was examined based on its potential impact on popularity scores, measured from 0-100 with 100 being the most popular.
+For further inference, the popularity score was then modeled using multiple linear regression, with the other variables as predictors.
+
+## Popularity
+
+First, we assess how popularity is distributed across our dataset using a histogram.
+
+```{r popularity-histogram}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Histogram of Popularity Scores
+#| fig-align: center
+
+ggplot(spotify_data, aes(x = popularity)) +
+  geom_histogram(bins = 14, fill = "blue", color = "black") + 
+  labs(title = "Histogram of Popularity",
+       x = "Popularity",
+       y = "Density") +
+  theme_minimal() + 
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5))
+
+```
+
+Against expectations, the histogram in [@popularity-histogram] shows no skewed distribution toward higher popularity scores for the most popular artists.
+Although most songs have scores between 54 and 86, which indicates relatively lasting popularity, the curve looks somewhat bell-shaped.
+This indicates that there is an even spread of top performing and mediocre artists, even within the subset of the best performing artists in the world.
+
+## Valence
+
+Next, we focus on valence.
+As one of Spotify's algorithmically generated metrics, valence attempts to measure the musical "positivity" of a song's audio, with scores ranging between 0.0 and 1.0.
+As Spotify's documentation says: "Tracks with high valence sound more positive (e.g.
+happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g.
+sad, depressed, angry)." [@spotifydocs].
+
+```{r scatter-valence}
 #| message: false
@@
 #| fig-cap: CAPTION
-#| fig-align: ALIGNMENT
+#| fig-align: center
+
+# Scatter plot of Popularity vs. Energy
+ggplot(spotify_data, aes(x = valence, y = popularity)) +
+  geom_point(alpha = 0.6, color = "coral") +
+  ggtitle("Popularity vs. Valence") +
+  xlab("Valence") +
+  ylab("Popularity Score") +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5)
+  )
 
@@
 
-### Valence
+When plotting valence values against popularity scores, [@scatter-valence] shows no clear linear trend between valence and popularity, although a clumping of high valence scores near 1.00, suggests potential data issues.
 
-```{r boilerplate-r-chunk}
+## Danceability
+
+Like valence, danceability is another algorithmically calculated metric by Spotify.
+From their API documentation, "Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
+A value of 0.0 is least danceable and 1.0 is most danceable.".
+
+```{r scatter-danceability}
 #| message: false
@@
 #| fig-cap: CAPTION
-#| fig-align: ALIGNMENT
+#| fig-align: center
+
+# Scatter plot of Popularity vs. Energy
+ggplot(spotify_data, aes(x = danceability, y = popularity)) +
+  geom_point(alpha = 0.6, color = "coral") +
+  ggtitle("Popularity vs. Danceability") +
+  xlab("Danceability") +
+  ylab("Popularity Score") +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5)
+  )
 
@@
 
-### Danceability
+Plotting the danceability scores against popularity in [@scatter-danceability] shows a weak association between danceability and popularity score.
 
-```{r boilerplate-r-chunk}
+## Mode
+
+Musical mode indicates whether a song is in a major ('happier') or minor key ('sadder').
+Unlike valence, which is calculated from a variety of physical metrics, musical mode is inherent to musical piece and can be inferred from the arrangement of notes in a song.
+In the dataset, 1 indicates the song is major while 0 is minor.
+
+```{r maj-min-barchart}
 #| message: false
@@
 #| fig-cap: CAPTION
-#| fig-align: ALIGNMENT
+#| fig-align: center
+
+# Summarizing the data to count explicit vs. non-explicit songs
+song_counts <- spotify_data |>
+  group_by(mode) |>
+  summarise(count = n()) |>
+  mutate(explicit = if_else(mode == 1, "Major", "Minor"))
+
+# Plotting the counts of explicit and non-explicit songs
+ggplot(song_counts, aes(x = mode, y = count, fill = explicit)) +
+  geom_col() +  # geom_col is used for bar plots with predefined data summaries
+  ggtitle("Count of Major vs. Minor Songs") +
+  xlab("Song Mode") +
+  ylab("Number of Songs") +
+  scale_fill_manual(values = c("Major" = "skyblue", "Minor" = "salmon")) +
+  theme_minimal() +  # Adds a minimal theme for better aesthetics
+  theme(
+    axis.text.x = element_blank(),  # Hides the x-axis text
+    axis.ticks.x = element_blank(), 
+    plot.title.position = "plot",
+    plot.title = element_text(hjust = 0.5)
+  )
 
@@
 
-### Explicit Status
+[@maj-min-barchart] shows that the most popular songs are in a major key, with more than twice as many major songs than minor songs.
 
-```{r boilerplate-r-chunk}
+```{r density-plot}
 #| message: false
@@
 #| fig-cap: CAPTION
-#| fig-align: ALIGNMENT
+#| fig-align: center
+
+# Comparing popularity distributions for different modes
+ggplot(spotify_data, aes(x = popularity, fill = as.factor(mode))) +
+  geom_density(alpha = 0.5) +
+  ggtitle("Popularity Distribution by Mode") +
+  xlab("Popularity Score") +
+  ylab("Density") +
+  scale_fill_manual(values = c("dodgerblue", "gold"), labels = c("Minor", "Major")) +
+  theme(
+    plot.title = element_text(hjust = 0.5)
+  )
 
@@
 
-### Loudness
+Creating a density plot comparison in [@density-plot] shows that although there are more major songs overall, there actually appears to be more songs in minor keys that have higher popularity scores.
+This means that audiences may tend to have a higher preference for songs in minor keys over major.
 
-```{r boilerplate-r-chunk}
+## Explicit Status
+
+Spotify records whether a song contains explicit lyrics through reporting by the music publishers.
+
+```{r explicit-bar-chart}
 #| message: false
@@
 
-### Duration
+[@explicit-bar-chart] shows an imbalanced amount of non-explicit songs among popular artists, with more than seven times that of the number of explicit songs.
+That could be a artifact of radio stations preferences for songs without explicit language.
 
-```{r boilerplate-r-chunk}
+```{r explicit-density-plot}
 #| message: false
@@
 #| fig-cap: CAPTION
-#| fig-align: ALIGNMENT
+#| fig-align: center
+
+# Comparing popularity distributions for explicit status
+ggplot(spotify_data, aes(x = popularity, fill = as.factor(explicit))) +
+  geom_density(alpha = 0.5) +
+  ggtitle("Popularity Distribution by Explicit Status") +
+  xlab("Popularity Score") +
+  ylab("Density") +
+  scale_fill_manual(values = c("dodgerblue", "gold"), labels = c("Not Explicit", "Explicit")) 
+  theme(
+    plot.title = element_text(hjust = 0.5)
+  )
 
@@
 
-# Model
+Comparing the distribution of popularity scores between explicit and not-explicit songs shows a different story when it comes to popularity score on Spotify.
+Here, we see that explicit songs have the highest density at higher popularity scores, which indicates that explicit songs tend to be given higher popularity scores on average.
+This could mean that explicit songs are simply more popular on streaming services like Spotify, where users can choose what music they would like to listen to.
 
-Create a Correl matrix to motivate choosing of variables.
+## Loudness
+
+The average loudness of a song, in decibels (dB) is calculated by averaging the height of the waveforms in a particular song.
+In [@scatter-loudness] we plot loudness against popularity to assess any relationship between the two variables.
+
+```{r scatter-loudness}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: center
+
+# Scatter plot of Popularity vs. Energy
+ggplot(spotify_data, aes(x = loudness, y = popularity)) +
+  geom_point(alpha = 0.6, color = "coral") +
+  ggtitle("Popularity vs. Loudness") +
+  xlab("Loudness") +
+  ylab("Popularity Score") +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5)
+  )
+
+```
+
+[@scatter-loudness] shows no clear relationship between loudness and popularity, likely due to the varied nature of popular music and other confounding factors.
+That makes sense, since there could be songs that are "known" for being soft and emotional, while others are loud.
+Given the varied nature of pop music, it is to be expected.
+
+## Duration
+
+Duration's effect on popularity was then investigated.
+
+```{r duration-histogram}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: center
+
+ggplot(spotify_data, aes(x = duration_secs/60)) +
+  geom_histogram(bins = 20, fill = "blue", color = "black") + 
+  labs(title = "Histogram of Song Duration (Mins)",
+       x = "Duration (mins)",
+       y = "Density") +
+  theme_minimal() + 
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5))
+
+```
+
+Most popular songs range from 2.5 to 5 minutes based on [@duration-histogram], but [@scatter-duration] reveals no linear relationship between duration and popularity.
+
+```{r scatter-duration}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: center
+
+# Scatter plot of Popularity vs. Energy
+ggplot(spotify_data, aes(x = duration_secs, y = popularity)) +
+  geom_point(alpha = 0.6, color = "coral") +
+  ggtitle("Popularity vs. Duration") +
+  xlab("Duration") +
+  ylab("Popularity Score") +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5)
+  )
+
+```
+
+# Data Modelling {#sec-model}
+
+While individual variable analyses provide some surface-level insights into the determinants of the popularity of a song, music's complexity necessitates considering variables together for reasonable and informative inference.
+
+## Model justification
+Creating a linear regression model is primarily motivated by the fact that it can account for relationships between the variables and the popularity, as well the relationships between the predictor variables themselves.
+Running a linear regression model allows us to isolate effects of a particular variable while also controlling for the effects of the other variables at the same time.
+
+Before conducting the analysis, however, several tests were conducted (shown in [@sec-testing]) to verify the data fit within the linear regression assumptions of linearity, normally distributed errors, and homoscedasticity of residuals.
+This ensured that the data was well suited for analysis using a multiple regression.
 
@@
 
-### Model justification
+We are interested in investigating the relationship between several variables and popularity.
+These variables are: valence, danceability, mode, explicit lyrics, loudness, and duration.
+For each of the variables we are investigating, we have a null and an alternative hypothesis.
 
-# Results
+**The null hypothesis ($H_0$):** there is no significant linear relationship between one of the variables and a given Spotify popularity score, holding everything else constant.
 
-I guess put the results of the model?
+**The alternative hypothesis ($H_1$):** a significant linear relationship exists between one of the variables and a given Spotify popularity score, holding everything else constant.
 
-# Discussion
+For a particular variable, a low p-value for the regression coefficient would provide evidence against the null hypothesis, indicating that the variable has a meaningful effect on Spotify popularity score.
+On the other hand, if the p-value is high, this would suggest that there is not enough evidence to reject the null hypothesis, and there may be no significant linear relationship between the variable and the Spotify popularity score.
+Ultimately, the results of the analysis will inform whether the null hypothesis can be rejected or not.
 
-What does everything mean?
+The equation for our linear model can be written as follows:
 
-## First discussion point {#sec-first-point}
+$S = \beta_0 + \beta_1V + \beta_2D + \beta_3M + \beta_4E + \beta_5L + \beta_6T + \epsilon$,
 
-## Second discussion point {#sec-second-point}
+where,
 
-## Third discussion point {#sec-third-point}
+- $S$ is a song's Spotify popularity score.
+
+- $\beta_0$ is the model constant, better known as $S$ when all the other variables are 0,
+
+- $\beta_1$ is effect of valence (measured 0-1) on $S$,
+
+- $\beta_2$ is effect of danceability (measured 0-1) on $S$,
+
+- $\beta_3$ is effect of musical mode (0 or 1) on $S$,
+
+- $\beta_4$ is effect of explicit lyrics (0 or 1) on $S$,
+
+- $\beta_5$ is effect of loudness (measured in decibels) on $S$,
+
+- $\beta_6$ is effect of song duration (measured in seconds) on $S$, and
+
+- $\epsilon$ is the the random error term, which accounts for variation in $S$ that is not explained by the relationship with any of the other variables.
+
+The aim of computing a regression is to estimate a line with the parameters $\beta_0$ to $\beta_6$ so that the difference (error) between the predicted line and the data points is minimized.
+By doing this, we get an equation for a line that best fits the data, allowing for the estimation of a given Spotify score for a set of audio features.
+
+In the linear model, the dependent variable $S$ represents the Spotify popularity score measured from 0 to 100.
+
+- The variable $V$ denotes the valence (emotional positivity) of the song, measured from 0 to 1, where 1 is the most positive.
+The coefficient $\beta_1$ quantifies how a one-unit increase in valence (e.g., from 0.5 to 1.0) affects the popularity score $S$.
+- The variable $D$ represents the danceability of the song, also scaled from 0 to 1, with higher values indicating more danceability.
+The coefficient $\beta_2$ captures the change in $S$ for a one-unit increase in danceability.
+
+- The binary variable $M$ indicates the musical mode (0 for minor, 1 for major), with $\beta_3$ reflecting the difference in popularity between major and minor keys.
+
+- The binary variable $E$ indicates explicit lyrics (0 for non-explicit, 1 for explicit), where $\beta_4$ estimates the effect of explicit content on popularity.
+
+- The variable $L$ denotes the loudness of the song in decibels, with $\beta_5$ quantifying how a one-decibel increase in loudness impacts the popularity score $S$.
+
+- Finally, $T$ represents the duration of the song in seconds, and $\beta_6$ estimates the effect of an additional second on the popularity score.
+
+# Results {#sec-results}
+
+By running our model with the data collected from Spotify, we get the following results:
+
+\newpage
+```{r model}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Load the model from an RDS file
+model <- readRDS("../models/model.rds")
+
+# Get summary
+model_summary <- summary(model)
+
+# Extract coefficients
+coefficients_df <- tibble(
+  Term = rownames(model_summary$coefficients),
+  Estimate = model_summary$coefficients[, "Estimate"],
+  `Std. Error` = model_summary$coefficients[, "Std. Error"],
+  Statistic = model_summary$coefficients[, "t value"],
+  `P-value` = model_summary$coefficients[, "Pr(>|t|)"]
+)
+
+kable(coefficients_df, 
+      digits = 2, 
+      caption = "Linear Model of Spotify Popularity Summary"
+    )
+```
+
+The intercept (65.85) represents the predicted value of the dependent variable when all the independent variables are set to zero.
+However, since most of the predictors are likely scaled or binary, the interpretation of the intercept is not very meaningful.
+
+Valence, which measures the emotional positivity of a song on a scale of 0 to 1, has a negative coefficient (-10.96).
+This suggests that, holding other variables constant, an increase in valence by one unit is associated with a decrease in the predicted Spotify popularity score by approximately 11 points.
+
+```{r effect-plot-valence}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: CAPTION
+#| fig-align: center
+
+effect_plot(model, 
+            pred = valence, 
+            interval = TRUE, 
+            plot.points = TRUE, 
+            jitter = 0.05
+          )
+
+```
+
+Danceability, also scaled from 0 to 1, has a positive coefficient (18.24).
+This indicates that, controlling for other factors, a one-unit increase in danceability is associated with an increase in the predicted popularity score by about 18 points.
+
+```{r effect-plot-danceability}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+effect_plot(model, 
+            pred = danceability, 
+            interval = TRUE, 
+            plot.points = TRUE, 
+            jitter = 0.05
+          )
+
+```
+
+Mode is a binary variable (likely 0 for minor and 1 for major keys), and its coefficient (-0.85) is not statistically significant (p-value = 0.37), suggesting that the musical mode may not have a significant effect on the popularity score after accounting for other predictors.
+
+```{r effect-plot-mode}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+effect_plot(model, 
+            pred = mode, 
+            interval = TRUE, 
+            plot.points = TRUE, 
+            jitter = 0.05
+          )
+
+```
+
+Explicit is another binary variable (possibly indicating the presence of explicit lyrics), with a positive coefficient (10.10).
+This implies that, holding other variables constant, songs with explicit lyrics are predicted to have a popularity score that is approximately 10 points higher than non-explicit songs.
+
+```{r effect-plot-explicit}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+effect_plot(model, 
+            pred = explicit, 
+            interval = TRUE, 
+            plot.points = TRUE, 
+            jitter = 0.05
+          )
+
+```
+
+Loudness, measured in decibels, has a positive coefficient (0.82), indicating that an increase in loudness by one decibel is associated with an increase in the predicted popularity score by 0.82 points, controlling for other factors.
+
+```{r effect-plot-loudness}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+effect_plot(model, 
+            pred = loudness, 
+            interval = TRUE, 
+            plot.points = TRUE, 
+            jitter = 0.05
+          )
+
+```
+
+Duration_secs, likely representing the song duration in seconds, has a positive coefficient (0.01) with a p-value of 0.10, which is marginally significant.
+This suggests that, holding other variables constant, an increase in song duration by one second is associated with a slight increase in the predicted popularity score by 0.01 points.
+
+```{r effect-plot-duration}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+effect_plot(model, 
+            pred = duration_secs, 
+            interval = TRUE, 
+            plot.points = TRUE, 
+            jitter = 0.05
+          )
+
+```
+
+Taken together, the regression results indicate that danceability, explicit lyrics, loudness, and, to a lesser extent, song duration are associated with higher Spotify popularity scores, while valence (emotional positivity) has a negative association.
+The musical mode does not appear to have a significant effect on popularity after accounting for other predictors.
+
+# Discussion {#sec-discussion}
+
+## Considerations of Spotify Metrics {#sec-first-point}
+
+## Impact of Results {#sec-second-point}
 
 ## Weaknesses and next steps {#sec-weaknesses-next}
-- Standard errors be huge. That means the regression is very imprecise. One way to
-resolve that would be to get a larger sample size. 
-- Sample size is kind of small - but that's the point? 
-Maybe constructing a fuller dataset with the entire discographies 
-of these people using Spotify's assistance would allow us to get more precise
-estimates of effects. 
-- People could repeat this experiment on a genre-based basis or using different 
-ratings of artists. Perhaps there are genre-specific relationships within other 
-variables in the spotify API that I haven't covered here.
+
+Although estimates are relatively precise, increasing the sample size of songs could help resolve this issue and improve precision.
+Additionally, the Billboard data used ends in 2015, limiting the analysis's relevance to more recent years.
+Obtaining an updated list from Billboard or another source would enhance the dataset's timeliness.
+While the sample size is relatively small, this aligns with the focus on top artists.
+
+Moving forward, constructing a more comprehensive dataset encompassing the entire discographies of these artists, with Spotify's assistance, could yield more precise effect estimates.
+Repeating this analysis on top artists within specific genres or using different artist rating criteria could uncover genre-specific relationships or highlight the influence of other Spotify API variables not covered in this analysis.
+Exploring alternative data sources or metrics beyond Spotify's popularity score could also provide additional insights into the determinants of a song's widespread and enduring appeal.
 
@@
 
-# Model Testing
+## Model Testing {#sec-testing}
+
+```{r model-testing}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Residuals Plot
+plot(model$fitted.values, residuals(model),
+     main = "Residuals vs Fitted",
+     xlab = "Fitted values",
+     ylab = "Residuals")
+abline(h = 0, col = "red")
+
+# Q-Q Plot
+qqnorm(residuals(model))
+qqline(residuals(model), col = "steelblue")
+
+#Scale-Location Plot 
+plot(model$fitted.values, sqrt(abs(residuals(model))),
+     main = "Scale-Location Plot",
+     xlab = "Fitted values",
+     ylab = "Square Root of |Residuals|")
+
+# Residual vs Leverage Plot
+plot(hatvalues(model), residuals(model),
+     main = "Residuals vs Leverage",
+     xlab = "Leverage",
+     ylab = "Residuals")
+abline(h = 0, col = "red")
+
+```
 

### 2024-04-23 scripts/00-simulate_data.R

@@ -11,13 +11,11 @@
 library(tidyverse)
 library(stringi)
 library(ggplot2)
-library(beepr)
 library(broom)
 library(broom.mixed)
 library(knitr)
 library(modelsummary)
 library(purrr)
-library(rstanarm)
 library(testthat)
 library(tidyverse)
 
@@ -82,13 +80,6 @@ print(range_checks)
 
 ## Simulate some graphs (because the API kicked me out lol)
 
-## Liveness
-ggplot(simulated_data, aes(x = liveness)) +
-  geom_histogram(bins = 45, fill = "blue", color = "black") + # You can adjust the number of bins as needed
-  labs(title = "Histogram of Liveness",
-       x = "Liveness",
-       y = "Density") +
-  theme_minimal()
 
 ## Popularity
 ggplot(simulated_data, aes(x = popularity)) +
@@ -114,9 +105,14 @@ ggplot(simulated_data, aes(x = explicit)) +
 
 # danceability 
 
-# instrumentalness
-
 # duration
+
+ggplot(simulated_data, aes(x = duration_secs)) +
+  geom_histogram(bins = 45, fill = "blue", color = "black") + # Adjust the number of bins as needed
+  labs(title = "Histogram of Popularity",
+       x = "Popularity",
+       y = "Density") +
+  theme_minimal()
 
 
 # Try doing a model

### 2024-04-24 paper/datasheet.qmd
styled the code. edited paper.

---
title: "Datasheet for 'A dataset'"
subtitle: "My subtitle if needed"
author: 
  - Luca Carnegie
thanks: "Code and data are available at: https://github.com/lcarnegie/popularity-modeling"
date: "April 24 2024"
abstract: "I construct a datasheet going over the details of the dataset used for my analysis in 'The Anatomy of a Hit'"
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebruetal.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable analysis of audio features on Spotify and their effect on a song's popularity score on Spoitfy. 
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Spotify created the data on the dataset. 
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The creation of the data was likely funded by Spotify as well. 
4. *Any other comments?*
    - No.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- TBD
2. *How many instances are there in total (of each type, if appropriate)?*
	- TBD
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- TBD
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- TBD
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- TBD
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- TBD
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- TBD
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- TBD
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- FILL OUT
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- TBD
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No. 
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No. 
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- No. 
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No. 
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No. 
16. *Any other comments?*
	- No.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- Some of the data (e.g. mode and explicit columns) were collected from 
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- They were collected using crowdsourced labour and machine listening techniques. 
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Luca Carnegie (the author of the paper) was the sole person involved in data collection. 
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- This data was collected over 1 hour by repeatedly querying the Spotify API for 
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	-  No, none were conducted. 
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	-  Not applicable
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	-  Not applicable
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Not applicable
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Not applicable
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Not applicable
12. *Any other comments?*
	- No

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- No. 
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes. It is available within this repository at popularity-modeling/data/raw_data. 
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- Yes. The data was processed using the RStudio environment, which is a freely available software package.
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- It was used to write the paper "The Anatomy of a Hit", which analyzes the determinants of popularity scores of songs on Spotify, using the discographies of top pop music artists. 
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- It can be accessed at https://github.com/lcarnegie/popularity-modeling 
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset will be hosted and distributed through the GitHub repository at https://github.com/lcarnegie/popularity-modeling
3. *When will the dataset be distributed?*
	- It is currently available for perusal and analysis, as of April 24. 2024
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- No. 
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No. However, through their API terms and conditions Spotify outlines that Spotify content may not be downloaded, that visual content be kept in its original form and that content attribution.

6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No. 
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The dataset will be hosted within this repository at: 
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- Luca Carnegie is contactable at luca.carnegie@mail.utoronto.ca
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No. 
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- It will likely not be updated after the end of writing this paper. 
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- Not applicable. 
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- If the dataset is updated, older versions will continue to be supported. 
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Luca Carnegie is heartily open for collaborations to extend and augment this dataset. He is available at luca.carnegie@mail.utoronto.ca 
8. *Any other comments?*
	- No.

\newpage

# References

### 2024-04-24 paper/paper.qmd
styled the code. edited paper.

@@
 
-Unlike other datasets like The Million Song dataset used by Pham et al., Spotify's web API provides ready-made quantitative audio features (tempo, key, danceability, etc.) as well as calculated popularity score based on recent and total play counts for artists' songs, facilitating analysis of popularity determinants.
+Unlike other datasets like The Million Song dataset used by Pham et al., Spotify's web API provides ready-made quantitative audio features (tempo, key, danceability, etc.) as well as calculated popularity score based on recent and total play counts for artists' songs, facilitating analysis of the determinants of a song's popularity.
 
-Since the popularity score is based on both recency of streams and total number of streams, continually high scores indicate frequent plays long after release.
-This means this score can be thought of measuring a song's enduring popularity.
+Since the Spotify popularity score is based on both recency of streams and total number of streams, continually high scores indicate frequent plays long after release.
+This score can then be thought of measuring a song's enduring popularity, whose qualities would be of interest to up and coming artists.
 As the leading U.S.
@@
 
-To analyze the audio feature determinants of popularity, I used R [@citeR], the tidyverse [@tidyverse] and related packages.
+To acquire, clean and analyze the data, I used R [@citeR], the tidyverse [@tidyverse] and related software packages.
 Billboard's "Greatest of All Time Hot 100 Artists" [@greatestofhot100] list identified popular artists.
-The audio features of the Billboard artists' songs and popularity data were then acquired from Spotify's API via the spotifyr [@spotifyr] package.
+The audio features of the Billboard artists' songs and popularity data were then downloaded from Spotify's API via the spotifyr [@spotifyr] package.
 Spotify data was current as of 2024, while Billboard rankings were from 2015.
@@
 
+<center>
 $S = \beta_0 + \beta_1V + \beta_2D + \beta_3M + \beta_4E + \beta_5L + \beta_6T + \epsilon$, 
+</center>
 
@@
 \newpage
+
 ```{r model}
@@
 )
-
 kable(coefficients_df, 
       digits = 2, 
+      format = "html",
       caption = "Linear Model of Spotify Popularity Summary"
@@
 The intercept (65.85) represents the predicted value of the dependent variable when all the independent variables are set to zero.
-However, since most of the predictors are likely scaled or binary, the interpretation of the intercept is not very meaningful.
+However, since most of the predictors are both scaled or binary, the interpretation of the intercept is not very meaningful in the context of this analysis.
 
@@
 
-Duration_secs, likely representing the song duration in seconds, has a positive coefficient (0.01) with a p-value of 0.10, which is marginally significant.
+Duration has a positive coefficient (0.01) with a p-value of 0.10, which is marginally significant.
 This suggests that, holding other variables constant, an increase in song duration by one second is associated with a slight increase in the predicted popularity score by 0.01 points.
@@
 
-## Considerations of Spotify Metrics {#sec-first-point}
+## Impact of Results {#sec-first-point}
 
-## Impact of Results {#sec-second-point}
+These findings offer valuable insights into the extensive efforts undertaken by music technology companies like Spotify to quantify and understand the multifaceted dynamics of music culture.
+While attempting to reduce the inherently complex artistic medium of music to quantitative measures may appear overly reductive or naive, our analysis demonstrates that even with relatively basic domain knowledge of the music industry, one can uncover meaningful trends and patterns in previously released and publicly consumed music.
+This realization speaks to the fundamental human propensity for discerning patterns, which extends to our preferences and consumption behaviors in domains as subjective as music, as evidenced by the popularity metrics on the Spotify platform.
+
+Furthermore, the analysis underscores the potential impact and utility of quantitative, data-driven methods in comprehending and potentially shaping cultural phenomena like music popularity.
+By developing a better understanding of the specific audio elements and characteristics associated with higher popularity and mainstream appeal, artists and music creators could strategically focus their creative efforts and artistic vision on incorporating these elements, potentially mitigating common challenges such as writer's block or creative stagnation.
 
@@
 
-Although estimates are relatively precise, increasing the sample size of songs could help resolve this issue and improve precision.
-Additionally, the Billboard data used ends in 2015, limiting the analysis's relevance to more recent years.
-Obtaining an updated list from Billboard or another source would enhance the dataset's timeliness.
-While the sample size is relatively small, this aligns with the focus on top artists.
+### Data and Measurement
 
-Moving forward, constructing a more comprehensive dataset encompassing the entire discographies of these artists, with Spotify's assistance, could yield more precise effect estimates.
-Repeating this analysis on top artists within specific genres or using different artist rating criteria could uncover genre-specific relationships or highlight the influence of other Spotify API variables not covered in this analysis.
-Exploring alternative data sources or metrics beyond Spotify's popularity score could also provide additional insights into the determinants of a song's widespread and enduring appeal.
+The age of the data used, with the Billboard Greatest of All Time list concluding in 2015, may not fully capture contemporary trends and shifts in music culture, especially given the rapid rise of social media platforms like TikTok that drive rapid virality and micro-trends in popular music consumption.
+Additionally, the crowdsourced nature of certain audio features like valence and danceability, initially derived from subjective assessments by college interns, introduces an element of human subjectivity and potential bias, which could undermine the validity and generalizability of the relationships observed between these features and popularity scores.
+The reliance solely on Spotify popularity scores as the dependent variable may overly represent newer music versus older releases that gained popularity through non-digital means.
+That said, the recency factor incorporated into Spotify's scoring algorithm could potentially mitigate this concern for tracks that have maintained enduring popularity over time, which gives this analysis merit.
+
+### Methodology
+
+Methodological limitations arising from the use of multiple regression complicate the interpreation of our results.
+Potential unknown multicollinearity, through having two audio features be correllated in some way, could affect the interpretation and relative influence of individual variables on popularity scores.
+Moreover, the use of a linear regression model, while providing good initial insights, could oversimplify the relationships by failing to capture non-linear or simultaneous interaction effects between the variables analyzed.
+Furthermore, the dataset's limited scope, comprising only the top 10 songs per artist and excluding some tracks due to missing audio feature data from Spotify's API, restricts the sample size and may impact the precision and robustness of the parameter estimates.
+
+### Next steps
+
+To address the weaknesses and expand upon the present findings, there are several avenues for future research emerge.
+Constructing a more comprehensive dataset with complete artist discographies and a broader array of audio features could better capture the complex relationships between musical characteristics and popularity.
+Exploring non-linear modeling techniques or machine learning algorithms may uncover nuanced patterns and interactions that linear models cannot detect.
+Incorporating alternative data sources or popularity metrics beyond the Spotify platform could provide further insights into the determinants of song appeal and resonance with audiences.
+Repeating the analysis on specific musical genres or using different lists of top artists could uncover genre-specific relationships or other variable influences that may be obscured in the present, more general analysis.
+Finally, validating and cross-referencing the results across multiple data sources would be crucial to assessing their reliability and generalizability to the whole pop music genre.
+
+Despite it's shortcomings, my approach affords merit as the first step by a cultural analyst in using the openly available data from Spotify to analyze and quantify what makes popular music popular.
+Through this analysis, I hope to inspire future work in attempting to understand the uniquely human elements of what makes our global culture tick.
 

### 2024-04-24 scripts/00-simulate_data.R

@@ -29,7 +29,7 @@ set.seed(2024)
 num_songs <- 100
 
 simulated_data <- tibble(
-  popularity = sample(0:100, num_songs, replace = TRUE), 
+  popularity = sample(0:100, num_songs, replace = TRUE),
   song = stri_rand_strings(num_songs, 10), # Random strings of length 10 for song titles
   artist = stri_rand_strings(num_songs, 7), # Random strings of length 7 for artist names
   energy = runif(num_songs, 0, 1), # Uniform distribution between 0 and 1
@@ -49,7 +49,7 @@ head(simulated_data)
 ## Test simulated Data using TestThat
 
 # 1. Summary Statistics Check
-summary_stats <- simulated_data |> 
+summary_stats <- simulated_data |>
   summarise(across(everything(), list(mean = mean, sd = sd, min = min, max = max)))
 print(summary_stats)
 
@@ -57,24 +57,26 @@ print(summary_stats)
 data_types <- sapply(simulated_data, class)
 print(data_types)
 
-# 3. Missing Values Test 
+# 3. Missing Values Test
 missing_vals <- sum(is.na(simulated_data))
 print(missing_vals)
 
 # 4. Uniqueness Test
-duplicated_songs <- simulated_data$song |> 
-                           duplicated() |> 
-                           sum()
+duplicated_songs <- simulated_data$song |>
+  duplicated() |>
+  sum()
 
-duplicated_artists <- simulated_data$artist |> 
-                               duplicated() |> 
-                               sum()
+duplicated_artists <- simulated_data$artist |>
+  duplicated() |>
+  sum()
 
 print(c(duplicated_songs, duplicated_artists))
 
 # 5. Range Validation Test
-range_checks <- sapply(simulated_data[c("energy", "valence", "danceability", "liveness", "instrumentalness")], 
-                       function(x) all(x >= 0 & x <= 1))
+range_checks <- sapply(
+  simulated_data[c("energy", "valence", "danceability", "liveness", "instrumentalness")],
+  function(x) all(x >= 0 & x <= 1)
+)
 print(range_checks)
 
 
@@ -84,61 +86,31 @@ print(range_checks)
 ## Popularity
 ggplot(simulated_data, aes(x = popularity)) +
   geom_histogram(bins = 45, fill = "blue", color = "black") + # Adjust the number of bins as needed
[…]

### 2024-04-24 scripts/01-download_data.R

@@ -39,12 +39,13 @@ artists <- trimws(gsub("/", " and ", names))
 
 # Take the list of artists and get data for each one
 
-# Initialize lists to store tibbles for each artist
+# Initialize lists to store the tibbles for each artist
 artist_audio_features_list <- list()
 artist_top_tracks_list <- list()
 
 # Loop through each artist and store their data
 for (artist_name in artists) {
+  
   # Get audio features
   audio_features <- get_artist_audio_features(artist_name)
   # Extract artist_id and use it to get top tracks

### 2024-04-24 scripts/03-test_data.R

@@ -4,7 +4,7 @@
 # Date: April 21 2023
 # Contact: luca.carnegie@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 01 -download_data.R and 02-data_cleaning.R
+# Pre-requisites: 01-download_data.R and 02-data_cleaning.R
 
 
 #### Workspace setup ####

### 2024-04-24 scripts/04-model.R

@@ -14,25 +14,20 @@ library(modelsummary)
 #### Read data ####
 analysis_data <- read_parquet("data/analysis_data/dataset.parquet")
 
-numeric_data <- analysis_data |> select(-song_id, -song_name, -artist_name)
-
-cor(numeric_data)
+numeric_data <- analysis_data |> select(-song_name, -artist_name)
 
 
 ### Model data ####
 
 model <-
   lm(
-    popularity ~  valence + danceability + explicit + mode + loudness + duration_secs,
-    data = analysis_data
+    popularity ~ valence + danceability + mode + explicit + loudness + duration_secs,
+    data = numeric_data
   )
 
-summary(sim_run_data_first_model)
 
 #### Save model ####
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  model,
+  file = "models/model.rds"
 )
-
-

### 2024-04-24 scripts/01-download_data.R
further styled code.

@@ -23,14 +23,14 @@ raw_data <-
     "https://www.billboard.com/charts/greatest-hot-100-artists/"
   )
 
-write_html(raw_data, 'data/raw_data/greatest100.html')
+write_html(raw_data, "data/raw_data/greatest100.html")
 
-raw_data <- read_html('data/raw_data/greatest100.html')
+raw_data <- read_html("data/raw_data/greatest100.html")
 
-names <- raw_data |> 
-          html_elements("h3") |> 
-          html_text() |> 
-          head(100)
+names <- raw_data |>
+  html_elements("h3") |>
+  html_text() |>
+  head(100)
 
 artists <- trimws(gsub("\\s+", " ", names))
 
@@ -45,20 +45,19 @@ artist_top_tracks_list <- list()
 
 # Loop through each artist and store their data
 for (artist_name in artists) {
-  
   # Get audio features
   audio_features <- get_artist_audio_features(artist_name)
   # Extract artist_id and use it to get top tracks
   artist_id <- audio_features$artist_id[1] # Assuming 'artist_id' is correctly named and not null
-  
+
   # Get top tracks for the US market
   top_tracks <- get_artist_top_tracks(artist_id, market = "US")
-  
+
   # Store data frames in respective lists
   artist_audio_features_list[[artist_name]] <- audio_features
   artist_top_tracks_list[[artist_name]] <- top_tracks
-  
-  Sys.sleep(sample(0:12)) #Stop for a random time to not get kicked out of API
+
+  Sys.sleep(sample(0:12)) # Stop for a random time to not get kicked out of API
   print(paste("Got", artist_name, "'s data"))
 }
 
@@ -75,5 +74,3 @@ top_tracks <- as_tibble(top_tracks_tibble_spotify)
 
 write_csv(audio_features, "data/raw_data/audio_features.csv")
 write_csv(top_tracks, "data/raw_data/top_tracks.csv")
-
-         

### 2024-04-24 scripts/02-data_cleaning.R

@@ -1,6 +1,6 @@
 #### Preamble ####
-# Purpose: Cleans the raw Spotify popularity/audio feature data and combines it 
-#          into a single dataset. 
+# Purpose: Cleans the raw Spotify popularity/audio feature data and combines it
+#          into a single dataset.
 # Author: Luca Carnegie
 # Date: 19 April 2024
 # Contact: luca.carnegie@mail.utoronto.ca
@@ -16,71 +16,72 @@ top_tracks <- read.csv("data/raw_data/top_tracks.csv")
 audio_features <- read.csv("data/raw_data/audio_features.csv")
 
 
-#Clean the two CSVs
-cleaned_top_tracks <- top_tracks |> 
-                      select(Artist_Name, name, popularity) |>
-                      rename(song_name = name) |>
-                      clean_names() |>
-                      mutate(song_id = paste(artist_name, song_name, sep = " - "))
-                      
-                      
-                          
-cleaned_audio_features <- audio_features|>
-                          mutate(explicit_ = if_else(explicit, 1, 0)) |>
-                          select(artist_name, 
-                                 track_name,
-                                 valence,
-                                 danceability,
-                                 explicit_,
-                                 loudness,
-                                 mode, 
-                                 duration_ms
-                                 ) |>
-                          rename(song_name = track_name) |>
-                          clean_names() |>
-                          mutate(song_id = paste(artist_name, song_name, sep = " - "))
-                          
+# Clean the two CSVs
+cleaned_top_tracks <- top_tracks |>
+  select(Artist_Name, name, popularity) |>
+  rename(song_name = name) |>
+  clean_names() |>
+  mutate(song_id = paste(artist_name, song_name, sep = " - "))
 
-#Merge based on song name
+
+
+cleaned_audio_features <- audio_features |>
+  mutate(explicit_ = if_else(explicit, 1, 0)) |>
+  select(
+    artist_name,
+    track_name,
+    valence,
+    danceability,
+    explicit_,
+    loudness,
+    mode,
+    duration_ms
+  ) |>
+  rename(song_name = track_name) |>
+  clean_names() |>
+  mutate(song_id = paste(artist_name, song_name, sep = " - "))
+
+
+# Merge based on song name
 cleaned_ds <- cleaned_audio_features |>
   # Perform an inner join on 'song_name' to filter only relevant songs from audio features
[…]

### 2024-04-24 paper/datasheet.qmd
finished datasheet

@@
 ---
-title: "Datasheet for 'A dataset'"
-subtitle: "My subtitle if needed"
+title: "Datasheet for Audio Features of Top 100 Greatest Billboard Artists Dataset"
 author: 
@@
 date: "April 24 2024"
-abstract: "I construct a datasheet going over the details of the dataset used for my analysis in 'The Anatomy of a Hit'"
+abstract: "I present a datasheet detailing the dataset used for my analysis in 'The Anatomy of a Hit'"
 format: pdf
@@
 
-Extract of the questions from @gebruetal.
+This datasheet is constructed in the image of an ideal datasheet written about by Gebru et.
+al in their paper "Datasheets for Datasets" by @gebruetal.
+I used the questions extracted from thier paper to motivate the discussion in my datasheet.
 
@@
 1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
-    - The dataset was created to enable analysis of audio features on Spotify and their effect on a song's popularity score on Spoitfy. 
+    - The dataset was created to enable analysis of audio features on Spotify and their effect on a song's popularity score on Spotify.
+
 2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
-    - Spotify created the data on the dataset. 
+    - Luca Carnegie, the author of the paper, created this dataset to write the paper "The Anatomy of a Hit". 
+    - The audio and popularity data itself was collected by Spotify, put into it's API and then collected by Luca to analyze for this paper.
+
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
-    - The creation of the data was likely funded by Spotify as well. 
+    - There was no funding involved with the creation of the dataset used for the paper.  
+   - Spotify was involved in the collection, processing and hosting of the data in the Spotify API, based on data from their own operations.
+
 4. *Any other comments?*
@@
 Please provide a description.*
-	- TBD
+	- Each instance represents a song.
+Each song is by an artist who was ranked in the Greatest Billboard Top 100 Artists, which is a list that compiled the top-performing artists in the Billboard Top 100, since it's inception in 1958.
+
 2. *How many instances are there in total (of each type, if appropriate)?*
-	- TBD
+   - There are 773 songs in the dataset.
+
 3.
@@
 If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
-	- TBD
+	- No. This dataset contains a sample of songs from the larger dataset of all artists/songs in the Spotify API. 
+	- This dataset was not intended to be representative of the larger dataset, as the point of it's creation was to analyze only the features of the best-performing Billboard artists, that are also on Spotify, with the goal of finding patterns within the audio features that were correlated with a high popularity score.
+
 4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
-	- TBD
+	- Each instance consists of a
+		- Artist Name
+		- Song Name 
+		- Popularity score (measured from 0-100) 
+		- Valence (musical positivity, measured from 0-1)
+		- Danceability (measured from 0-1)
+		- Mode (major or minor)
+		- Explicit (yes or no)
+		- Loudness (measured in decibels, dB)
+		- Duration (measured in seconds)
+
 5. *Is there a label or target associated with each instance? If so, please provide a description.*
-	- TBD
+	  - Each song is labelled by the artist name and song name.
+
 6.
@@
 This does not include intentionally removed information, but might include, for example, redacted text.*
-	- TBD
+	- No. This was checked and tested using the R programming environment.
+
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
-	- TBD
+	- Yes.
+Songs by the same artists are explicitly related by possessing the same artist name within their respective instances.
+If two individual artists made independent covers of the same songs, both versions would be differentiated by the instance's artist name rather than the song itself.
+
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
-	- TBD
+	- No.
+
 9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
-	- FILL OUT
+	- Possibly.
+Some of Spotify's metrics, such as valence and danceability are known to be subjective in their collection methods.
+Conversations by other researchers with Spotify engineers revealed that danceability scores were collected by asking college interns at a music intelligence company to rate each of the songs.
+From this, other calculated metrics like the valence could also be beholden to errors stemming from the subjectivity of the person making the rating.
+
 10.
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- TBD
+
+	- The dataset is self-contained, but as popularity of different artists changes quickly it is not expected to remain updated for very long, as Spotify updates their popularity data on a near-daily basis.
+
 11.
@@
 	- No.
+
 12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
 	- No.
+
 13.
@@
 	- No.
+
 14.
@@
 If so, please describe how.*
-	- No. 
+	- You can identify artists by their song name, but
+
 15.
@@
 	- No.
+
 16. *Any other comments?*
@@
 If so, please describe how.*
-	- Some of the data (e.g. mode and explicit columns) were collected from 
+	- Artist Name: taken from music publisher
+	- Song Name: taken from music publisher
+	- Popularity score: 
+	- Valence (musical positivity): 
+	- Danceability: before being acquired by Spotify in 2014, a music intelligence startup used labour by college interns, who rated songs as danceable or not.
+However, it is unclear if human rating still continues or whether that has been replaced by machine listening techniques.
+	- Mode (major or minor): machine listening techniques tagged songs as being in a major or minor key. 
+	- Explicit: machine listening techniques tagged songs as being explicit or not. 
+	- Loudness: calculated from taking the average loudness throughout the entire song
+	- Duration: taken from length of the recording of the song
+
 2.
@@
 How were these mechanisms or procedures validated?*
-	- They were collected using crowdsourced labour and machine listening techniques. 
+	- The data was collected using a combination of manual human curation and machine listening techniques, although specific details on the process were not able to be found due to the proprietary nature of Spotify's technology.
+
 3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
-	- TBD
+	- The sampling strategy was mainly based on convenience sampling - I chose only songs whose artists that appeared on the Billboard Hot 100 Greatest of All Time list. 
+	-It is of note that using the data collection process, certain songs appear on artists' top ten list did not have audio features that were accessible through the API, so they had to be discounted.
+As a consequence, there is an uneven amount of songs represented per artist.
+Therefore, certain artists may be more overrepresented than others.
+
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- Luca Carnegie (the author of the paper) was the sole person involved in data collection. 
+	- College interns were involved in parts of the collection process for variables like danceability, according to Drouillard.
+However, it is unclear is they were paid or not for their time.
+It is not clear if any of the songs present in the dataset were subject to this collection technique.
+It could be that some were rated through machine listening techniques and some through manual human rating.
+It is also unclear how the machine listening techniques were tuned.
+
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- This data was collected over 1 hour by repeatedly querying the Spotify API for 
+	- The timeframe for which the data on the Spotify API was collected is unclear.  
+	- For this analysis, however, the data was collected over 1 hour by querying the Spotify API for each artist, the audio features of their songs, and their top 10 tracks and storing them within a .csv file.
+
 6.
@@
 	- No, none were conducted.
+
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	-  Not applicable
+	- Not applicable, as the data concerns itself with songs and not data about individuals.
+
 8.
@@
 	- Not applicable
+
 9.
@@
 	- Not applicable
+
 10.
@@
 	- Not applicable
+
 11.
@@
 If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- Not applicable
+	- Not applicable.
+
 12. *Any other comments?*
-	- No
+	- No.
 
@@
 	- No.
+
 2.
@@
 	- Yes. It is available within this repository at popularity-modeling/data/raw_data.
+
 3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- Yes. The data was processed using the RStudio environment, which is a freely available software package.
+	- Yes. The data was processed using the RStudio environment, which is a freely available and downloaded software package searchable online.
+
 4. *Any other comments?*
@@
 	- It was used to write the paper "The Anatomy of a Hit", which analyzes the determinants of popularity scores of songs on Spotify, using the discographies of top pop music artists.
+
 2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
 	- It can be accessed at https://github.com/lcarnegie/popularity-modeling
+
 3. *What (other) tasks could the dataset be used for?*
-	- TBD
+	- It could be used, among other things, to categorize songs based on their characteristics, (e.g. genre, by popularity, etc.)
+
 4.
@@
 Is there anything a dataset consumer could do to mitigate these risks or harms?*
-	- TBD
-5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
+	- Yes.
+Since popularity data is by nature quite volatile, this dataset is expected to fall out of date very quickly.
+This dataset should only be used as a guide to motivate more sophisticated data collection and analysis methods that update as new data is collected.
+
+5. *Are there tasks for which the dataset should not be used? If so, please provide a description*
+	- Ideally, this dataset should not be used to cause evil or harm in the world.
+At the time of writing, no immediately dangerous or immoral acts are considered in this.
+It is my hope that those, if anyone, who uses it ultimately decides to do good with it in the hope of serving rather than hurting humanity.
+
 6. *Any other comments?*
-	- TBD
+	- No.
 
@@
 If so, please provide a description.*
-	- TBD
+	- At this stage, there is no plan to do such a thing.
+
 2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
 	- The dataset will be hosted and distributed through the GitHub repository at https://github.com/lcarnegie/popularity-modeling
+
 3. *When will the dataset be distributed?*
 	- It is currently available for perusal and analysis, as of April 24. 2024
+
 4.
@@
 	- No.
+
 5.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
-	- No. However, through their API terms and conditions Spotify outlines that Spotify content may not be downloaded, that visual content be kept in its original form and that content attribution.
+	- No.
+However, through their API terms and conditions Spotify outlines that Spotify content (particularly song recordings, rather than audio data) may not be downloaded, that visual content be kept in its original form and that content attribution.
 
@@
 	- No.
+
 7. *Any other comments?*
-	- TBD
+	- No.
 
@@
 1. *Who will be supporting/hosting/maintaining the dataset?*
-	- The dataset will be hosted within this repository at: 
+	- The dataset will be hosted within a GitHub repository at: https://github.com/lcarnegie/popularity-modeling
+
 2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
 	- Luca Carnegie is contactable at luca.carnegie@mail.utoronto.ca
+
 3. *Is there an erratum? If so, please provide a link or other access point.*
 	- No.
+
 4.
@@
 	- It will likely not be updated after the end of writing this paper.
+
 5.
@@
 	- Not applicable.
+
 6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
 	- If the dataset is updated, older versions will continue to be supported.
+
 7.
@@
 If so, please provide a description.*
-	- Luca Carnegie is heartily open for collaborations to extend and augment this dataset. He is available at luca.carnegie@mail.utoronto.ca 
+	- Luca Carnegie is heartily open for collaborations to extend and augment this dataset and it's goals. He is contactable at luca.carnegie@mail.utoronto.ca
+
 8. *Any other comments?*

### 2024-04-24 paper/paper.qmd
finished datasheet
the paper looks presentable!

@@
 
-Throughout history, musicians have aspired to write hit songs and achieve mainstream success in the music industry.
+Since the advent of recorded music, musicians have aspired to write hit songs and achieve mainstream success in the music industry.
 However, for major music companies, signing artists without a proven track record is a substantial risk, as most aspirants fail to produce hits.
-This has led to a concentrated market dominated by a small group of "superstars", such as Taylor Swift, who consistently top the charts [@rosen1981].
+This has led to a concentrated market dominated by a small group of "superstars", such as Taylor Swift or Ed Sheeran, who consistently top the charts [@rosen1981].
 Breaking into the music industry, then, presents a unique challenge for aspiring musicians' art: what elements contribute to a pop song's mass appeal?
 
-The field of Hit Song Science, which employs statistical methods to predict song popularity, has gained traction with the increasingly widespread availability of economical computing resources.
-Researchers [@dodds2010], students [@phametal2015], and likely record companies have attempted to construct models that attempt to explain song popularity, though the latter's efforts remain proprietary, leaving artists without access to this knowledge.
+The field of Hit Song Science, which employs statistical methods to predict song popularity, has gained traction in academia and industry with the increasingly widespread availability of economical computing resources.
+Researchers [@kim2021], students [@phametal2015], and likely record companies have attempted to construct models that attempt to explain song popularity, though the latter's efforts remain proprietary, leaving artists without access to knowledge that could prove valuable for advancing their careers.
 
 This study diverges from previous approaches by constructing a dataset solely from the discographies of the best-performing artists on the Billboard Hot 100, the preeminent metric of success in the United States music industry.
-By focusing exclusively on the most successful artists' music, the elements that define enduring hits in the US market can be more easily identified.
-The analysis employs multivariate regression to estimate a song's Spotify-generated popularity score, using several audio features derived from Spotify's API as predictors.
+I employ multivariate regression to estimate a song's Spotify-generated popularity score, using several audio features derived from Spotify's API as predictors.
+By focusing exclusively on the most successful artists' music, the elements that define enduring hits in the US market can be identified.
 
@@
 
-The paper is structured as follows: [@sec-data] provides a detailed overview of the dataset and analyzes each predictor individually; [@sec-model] outlines the regression model; [@sec-results] presents and discusses the results; and [@sec-discussion] critically examines the findings and their implications.
+The paper is structured as follows: [@sec-data] provides a detailed overview of the dataset and analyzes each predictor individually; [@sec-model] outlines the regression model; [@sec-results] presents and discusses the model's results; and [@sec-discussion] critically examines the findings and their implications.
 
@@
 
-Since the Spotify popularity score is based on both recency of streams and total number of streams, continually high scores indicate frequent plays long after release.
-This score can then be thought of measuring a song's enduring popularity, whose qualities would be of interest to up and coming artists.
-As the leading U.S.
-music streamer, Spotify's metrics can reasonably align with Billboard rankings, providing a relatively robust measure of a song's lasting popularity, which is crucial in attempting to infer the elements that make up a "hit".
+Since the Spotify popularity score is based on both recency of streams and total number of streams, continually high scores should indicate frequent plays long after release.
+Therefore, the score can be thought of measuring a song's enduring popularity, whose qualities would be of definite interest to up-and-coming artists.
+As the leading music streamer in the U.S., Spotify's metrics can reasonably align with Billboard rankings, providing a relatively robust measure of a song's lasting popularity, which is crucial in attempting to infer the elements that make up a "hit".
 
@@
 
-```{r tbl-sample}
+```{r tbl-sample, fig.height=3}
 #| message: false
@@
 #| warning: false
-#| tbl-cap: Popularity Scores and Audio Features
+#| tbl-cap: Sample of Dataset
 #| fig-align: center
@@
 
-spotify_data <- read_parquet("../data/analysis_data/dataset.parquet")
-
-sample_table <- spotify_data |>
+labelled_data <- spotify_data |>
   rename(
@@
     `Explicit` = explicit, 
-    `Loudness (dB)` = loudness, 
-    `Duration (seconds)` = duration_secs
+    `Loudness` = loudness, 
+    `Duration` = duration_secs
   ) |>
+  mutate(`Song Name` = "...") |> 
   head(5) |>
@@
 
-  kable(sample_table, format = "markdown") |>
-  kable_styling(full_width = F, font_size = 1) |>
-  column_spec(2, width = "1cm")  # Control the width of specific columns
+  kable(labelled_data, format = "markdown") |>
+  kable_styling(font_size = 10)  # Control the width of specific columns
 ```
 
-The primarily categorical dataset was first analyzed variable-by-variable.
+The primarily categorical dataset was then analyzed variable-by-variable.
 Each variable was examined based on its potential impact on popularity scores, measured from 0-100 with 100 being the most popular.
@@
 
-```{r popularity-histogram}
+```{r fig-pop-hist}
 #| message: false
@@
 
-Against expectations, the histogram in [@popularity-histogram] shows no skewed distribution toward higher popularity scores for the most popular artists.
+Against expectations, the histogram in @fig-pop-hist shows no skewed distribution toward higher popularity scores for the most popular artists.
 Although most songs have scores between 54 and 86, which indicates relatively lasting popularity, the curve looks somewhat bell-shaped.
@@
 
-```{r scatter-valence}
+```{r fig-scatter-val}
 #| message: false
@@
 #| warning: false
-#| fig-cap: CAPTION
+#| fig-cap: Plot of Popularity against Valence
 #| fig-align: center
@@
   )
-
 ```
 
-When plotting valence values against popularity scores, [@scatter-valence] shows no clear linear trend between valence and popularity, although a clumping of high valence scores near 1.00, suggests potential data issues.
+When plotting valence values against popularity scores, though, @fig-scatter-val shows no clear linear trend between valence and popularity, although a clumping of high valence scores near 1.00, suggests potential data issues.
 
@@
 Like valence, danceability is another algorithmically calculated metric by Spotify.
-From their API documentation, "Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
-A value of 0.0 is least danceable and 1.0 is most danceable.".
+In their API documentation, Spotify says that "Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
+A value of 0.0 is least danceable and 1.0 is most danceable."
 
-```{r scatter-danceability}
+```{r fig-scatter-danceability}
 #| message: false
@@
 #| warning: false
-#| fig-cap: CAPTION
+#| fig-cap: Plot of Popularity against Danceability
 #| fig-align: center
@@
 
-Plotting the danceability scores against popularity in [@scatter-danceability] shows a weak association between danceability and popularity score.
+Plotting the danceability scores against popularity in @fig-scatter-danceability shows a weak association between danceability and popularity score, however.
 
@@
 
-```{r maj-min-barchart}
+```{r fig-majmin-barchart, fig.height=3}
 #| message: false
@@
 #| warning: false
-#| fig-cap: CAPTION
+#| fig-cap: Counts of Major and Minor Songs
 #| fig-align: center
 
-# Summarizing the data to count explicit vs. non-explicit songs
+# Summarizing the data to count major vs. minor songs
 song_counts <- spotify_data |>
@@
   summarise(count = n()) |>
-  mutate(explicit = if_else(mode == 1, "Major", "Minor"))
+  mutate(mode = if_else(mode == 1, "Major", "Minor"))
 
-# Plotting the counts of explicit and non-explicit songs
-ggplot(song_counts, aes(x = mode, y = count, fill = explicit)) +
-  geom_col() +  # geom_col is used for bar plots with predefined data summaries
+# Plotting the counts of major and minor songs
+ggplot(song_counts, aes(x = mode, y = count, fill = mode)) +
+  geom_col() +
   ggtitle("Count of Major vs. Minor Songs") +
@@
   ylab("Number of Songs") +
-  scale_fill_manual(values = c("Major" = "skyblue", "Minor" = "salmon")) +
+  scale_fill_manual(
+    values = c("Major" = "skyblue", "Minor" = "salmon"),
+    name = "Musical Mode"
+    ) +
   theme_minimal() +  # Adds a minimal theme for better aesthetics
@@
 
-[@maj-min-barchart] shows that the most popular songs are in a major key, with more than twice as many major songs than minor songs.
+@fig-majmin-barchart shows that the most popular songs are in a major key, with more than twice as many major songs than minor songs.
 
-```{r density-plot}
+```{r fig-mode-density-plot}
 #| message: false
@@
 #| warning: false
-#| fig-cap: CAPTION
+#| fig-cap: Distribution of Popularity, by mode
 #| fig-align: center
@@
   ylab("Density") +
-  scale_fill_manual(values = c("dodgerblue", "gold"), labels = c("Minor", "Major")) +
+  scale_fill_manual(
+    values = c("dodgerblue", "gold"), labels = c("Minor", "Major"), 
+    name = "Musical Mode", 
+    ) +
   theme(
@@
 
-Creating a density plot comparison in [@density-plot] shows that although there are more major songs overall, there actually appears to be more songs in minor keys that have higher popularity scores.
+The density plot in @fig-mode-density-plot shows that although there are more major songs overall, there actually appears to be more songs in minor keys that have higher popularity scores.
 This means that audiences may tend to have a higher preference for songs in minor keys over major.
@@
 
-```{r explicit-bar-chart}
+```{r fig-explicit-bar}
 #| message: false
@@
 #| warning: false
-#| fig-cap: CAPTION
+#| fig-cap: Count of Explicit and Non-Explicit Songs
 #| fig-align: center
@@
   ylab("Number of Songs") +
-  scale_fill_manual(values = c("Not Explicit" = "skyblue", "Explicit" = "salmon")) +
+  scale_fill_manual(
+    values = c("Not Explicit" = "skyblue", "Explicit" = "salmon"),
+    name = "Explicit Content"
+    ) +
   theme_minimal()  # Adds a minimal theme for better aesthetics
-
 ```
 
-[@explicit-bar-chart] shows an imbalanced amount of non-explicit songs among popular artists, with more than seven times that of the number of explicit songs.
+@fig-explicit-bar shows an imbalanced amount of non-explicit songs among popular artists, with more than seven times that of the number of explicit songs.
 That could be a artifact of radio stations preferences for songs without explicit language.
 
-```{r explicit-density-plot}
+```{r fig-explicit-density}
 #| message: false
@@
 #| warning: false
-#| fig-cap: CAPTION
+#| fig-cap: Distribution of Popularity, by presence of explicit lyrics
 #| fig-align: center
@@
   ylab("Density") +
-  scale_fill_manual(values = c("dodgerblue", "gold"), labels = c("Not Explicit", "Explicit")) 
+  scale_fill_manual(
+    values = c("dodgerblue", "gold"), 
+    labels = c("Not Explicit", "Explicit"), 
+    name = "Explicit Content"
+    ) + 
   theme(
@@
   )
-
 ```
 
-Comparing the distribution of popularity scores between explicit and not-explicit songs shows a different story when it comes to popularity score on Spotify.
+Comparing the distribution of popularity scores between explicit and not-explicit songs in @fig-explicit-density shows a different story when it comes to popularity score on Spotify.
 Here, we see that explicit songs have the highest density at higher popularity scores, which indicates that explicit songs tend to be given higher popularity scores on average.
@@
 The average loudness of a song, in decibels (dB) is calculated by averaging the height of the waveforms in a particular song.
-In [@scatter-loudness] we plot loudness against popularity to assess any relationship between the two variables.
+In @fig-scatter-loudness we plot loudness against popularity to assess any relationship between the two variables.
 
-```{r scatter-loudness}
+```{r fig-scatter-loudness}
 #| message: false
@@
 #| warning: false
-#| fig-cap: CAPTION
+#| fig-cap: Plot of Popularity against Loudness
 #| fig-align: center
@@
   )
-
 ```
 
-[@scatter-loudness] shows no clear relationship between loudness and popularity, likely due to the varied nature of popular music and other confounding factors.
+The plot shows no clear linear relationship between loudness and popularity, likely due to the varied nature of popular music and other confounding factors.
 That makes sense, since there could be songs that are "known" for being soft and emotional, while others are loud.
@@
 Duration's effect on popularity was then investigated.
+According to @fig-duration-hist, most songs written by artists on the "Greatest Billboard Hot 100 Artists of All Time" range from 2.5 to 5 minutes.
+However, @fig-dur-scatter reveals no discernible linear relationship between song duration and Spotify popularity score.
 
-```{r duration-histogram}
+```{r fig-duration-hist}
 #| message: false
@@
 #| warning: false
-#| fig-cap: CAPTION
+#| fig-cap: Histogram of Song Durations
 #| fig-align: center
@@
 
-Most popular songs range from 2.5 to 5 minutes based on [@duration-histogram], but [@scatter-duration] reveals no linear relationship between duration and popularity.
-
-```{r scatter-duration}
+```{r fig-dur-scatter}
 #| message: false
@@
 #| warning: false
-#| fig-cap: CAPTION
+#| fig-cap: Plot of Popularity against Duration
 #| fig-align: center
@@
   )
-
 ```
@@
 For a particular variable, a low p-value for the regression coefficient would provide evidence against the null hypothesis, indicating that the variable has a meaningful effect on Spotify popularity score.
-On the other hand, if the p-value is high, this would suggest that there is not enough evidence to reject the null hypothesis, and there may be no significant linear relationship between the variable and the Spotify popularity score.
-Ultimately, the results of the analysis will inform whether the null hypothesis can be rejected or not.
+On the other hand, if the p-value is high for a given variable, this would suggest that there is not enough evidence to reject the null hypothesis, and there may be no significant linear relationship between that variable and the Spotify popularity score.
+Ultimately, the results of the analysis will inform whether there is any significant linear relationship, controlling for the other variables,between each the variables and a song's Spotify popularity score.
+
+The aim of computing a regression model is to estimate a line with the coefficients $\beta_0$ to $\beta_6$ so that the difference (error) between the predicted line and the data points is minimized.
+By doing this, we get an equation for a line that best fits the data, allowing for the estimation of a Spotify score for a given set of audio features.
 
@@
 
-<center>
-$S = \beta_0 + \beta_1V + \beta_2D + \beta_3M + \beta_4E + \beta_5L + \beta_6T + \epsilon$, 
-</center>
+\begin{center}
+$S = \beta_0 + \beta_1V + \beta_2D + \beta_3M + \beta_4E + \beta_5L + \beta_6T + \epsilon$
+\end{center}
 
-where,
-
-- $S$ is a song's Spotify popularity score.
-
-- $\beta_0$ is the model constant, better known as $S$ when all the other variables are 0,
-
-- $\beta_1$ is effect of valence (measured 0-1) on $S$,
-
-- $\beta_2$ is effect of danceability (measured 0-1) on $S$,
-
-- $\beta_3$ is effect of musical mode (0 or 1) on $S$,
-
-- $\beta_4$ is effect of explicit lyrics (0 or 1) on $S$,
-
-- $\beta_5$ is effect of loudness (measured in decibels) on $S$,
-
-- $\beta_6$ is effect of song duration (measured in seconds) on $S$, and
-
-- $\epsilon$ is the the random error term, which accounts for variation in $S$ that is not explained by the relationship with any of the other variables.
-
-The aim of computing a regression is to estimate a line with the parameters $\beta_0$ to $\beta_6$ so that the difference (error) between the predicted line and the data points is minimized.
-By doing this, we get an equation for a line that best fits the data, allowing for the estimation of a given Spotify score for a set of audio features.
-
-In the linear model, the dependent variable $S$ represents the Spotify popularity score measured from 0 to 100.
+The dependent variable $S$ represents the Spotify popularity score measured from 0 to 100.
 
@@
 The coefficient $\beta_1$ quantifies how a one-unit increase in valence (e.g., from 0.5 to 1.0) affects the popularity score $S$.
+
 - The variable $D$ represents the danceability of the song, also scaled from 0 to 1, with higher values indicating more danceability.
@@
 
-- Finally, $T$ represents the duration of the song in seconds, and $\beta_6$ estimates the effect of an additional second on the popularity score.
+- Finally, $T$ (for time), represents the duration of the song in seconds, and $\beta_6$ estimates the effect of an additional second on the popularity score.
 
@@
 
-By running our model with the data collected from Spotify, we get the following results:
+By running our model with the data collected from Spotify, we get the  results in @tbl-model:
 
-\newpage
-
-```{r model}
+```{r tbl-model}
 #| message: false
@@
 
-Valence, which measures the emotional positivity of a song on a scale of 0 to 1, has a negative coefficient (-10.96).
-This suggests that, holding other variables constant, an increase in valence by one unit is associated with a decrease in the predicted Spotify popularity score by approximately 11 points.
-
-```{r effect-plot-valence}
+```{r effect-plot-valence, fig.height=3}
 #| message: false
@@
 #| warning: false
-#| fig-cap: CAPTION
+#| fig-cap: Effect Plot of Valence on Popularity
 #| fig-align: center
@@
 
-Danceability, also scaled from 0 to 1, has a positive coefficient (18.24).
-This indicates that, controlling for other factors, a one-unit increase in danceability is associated with an increase in the predicted popularity score by about 18 points.
+Valence, which measures the emotional positivity of a song on a scale of 0 to 1, has a negative coefficient (-10.96).
+This suggests that, holding other variables constant, an increase in valence by one unit is associated with a decrease in the predicted Spotify popularity score by approximately 11 points.
 
-```{r effect-plot-danceability}
+```{r effect-plot-danceability, fig.height=3}
 #| message: false
@@
 #| warning: false
+#| fig-caption: Effect Plot of Danceability on Popularity
 #| fig-align: center
@@
           )
-
 ```
 
-Mode is a binary variable (likely 0 for minor and 1 for major keys), and its coefficient (-0.85) is not statistically significant (p-value = 0.37), suggesting that the musical mode may not have a significant effect on the popularity score after accounting for other predictors.
+Danceability, also scaled from 0 to 1, has a positive coefficient (18.24).
+This indicates that, controlling for other factors, a one-unit increase in danceability is associated with an increase in the predicted popularity score by about 18 points.
 
-```{r effect-plot-mode}
+```{r effect-plot-mode, fig.height=3}
 #| message: false
@@
 
-```{r effect-plot-explicit}
+```{r effect-plot-explicit, fig.height=3}
 #| message: false
@@
 
-```{r effect-plot-loudness}
+```{r effect-plot-loudness, fig.height=3}
 #| message: false
@@
 
-```{r effect-plot-duration}
+```{r effect-plot-duration, fig.height=3}
 #| message: false
@@
 
-## Impact of Results {#sec-first-point}
+These findings offer valuable insights into the extensive efforts undertaken by music technology companies like Spotify to quantify and understand the multifaceted dynamics of music culture.
+The analysis underscores the potential impact and utility of quantitative, data-driven methods in comprehending and potentially shaping cultural phenomena like popular music.
 
-These findings offer valuable insights into the extensive efforts undertaken by music technology companies like Spotify to quantify and understand the multifaceted dynamics of music culture.
-While attempting to reduce the inherently complex artistic medium of music to quantitative measures may appear overly reductive or naive, our analysis demonstrates that even with relatively basic domain knowledge of the music industry, one can uncover meaningful trends and patterns in previously released and publicly consumed music.
-This realization speaks to the fundamental human propensity for discerning patterns, which extends to our preferences and consumption behaviors in domains as subjective as music, as evidenced by the popularity metrics on the Spotify platform.
+## Impact of Results
 
-Furthermore, the analysis underscores the potential impact and utility of quantitative, data-driven methods in comprehending and potentially shaping cultural phenomena like music popularity.
-By developing a better understanding of the specific audio elements and characteristics associated with higher popularity and mainstream appeal, artists and music creators could strategically focus their creative efforts and artistic vision on incorporating these elements, potentially mitigating common challenges such as writer's block or creative stagnation.
+While attempting to reduce the inherently complex artistic medium of music to quantitative measures can in some cases be naïve and reductive, having a data-driven understanding of what elements made the most popular artists who they are could have crucial implications for the life and work of up and coming popular artists.
+The analysis demonstrates that even with relatively basic domain knowledge of the music industry, anyone can uncover interesting trends and patterns in previously released music.
 
-## Weaknesses and next steps {#sec-weaknesses-next}
+Armed with the knowledge of what elements work in making music popular, artists and music creators can be more empowered in their creative process.
+Using the insights from my analysis, they can be enabled to strategically focus their creative efforts and artistic vision on incorporating these elements, potentially mitigating common challenges such as writer's block or creative stagnation.
+They could also choose to diverge from these elements in order to explore new genres or music styles that may not be popular now, but could be in the future.
+Knowing the current trends now could inspire artists to make better art that effectively and innovatively builds on what has come before.
 
-### Data and Measurement
+## Weaknesses
 
+There are weaknesses and limitations in my research that warrant consideration.
 The age of the data used, with the Billboard Greatest of All Time list concluding in 2015, may not fully capture contemporary trends and shifts in music culture, especially given the rapid rise of social media platforms like TikTok that drive rapid virality and micro-trends in popular music consumption.
+
 Additionally, the crowdsourced nature of certain audio features like valence and danceability, initially derived from subjective assessments by college interns, introduces an element of human subjectivity and potential bias, which could undermine the validity and generalizability of the relationships observed between these features and popularity scores.
-The reliance solely on Spotify popularity scores as the dependent variable may overly represent newer music versus older releases that gained popularity through non-digital means.
-That said, the recency factor incorporated into Spotify's scoring algorithm could potentially mitigate this concern for tracks that have maintained enduring popularity over time, which gives this analysis merit.
+The reliance solely on Spotify popularity scores as the dependent variable could bias the results toward newer music versus older, more established releases, although the recency factor incorporated into Spotify's scoring algorithm could potentially mitigate this concern for tracks that have maintained enduring popularity over time.
+Since the mechanics of calculating the Spotify popularity score are proprietary, there is
 
-### Methodology
+There could also be potential methodological issues stemming from the complex and interrelated nature of the data itself.
+Most obviously, potential multicollinearity among the predictor variables (audio features) could affect the interpretation and relative influence of individual variables on popularity scores.
+In particular, given the little amount of reliable information about how a Spotify popularity score is calculated, it is worth questioning whether if certain algorithmically generated scores such as danceability are associated with popularity scores by design.
 
-Methodological limitations arising from the use of multiple regression complicate the interpreation of our results.
-Potential unknown multicollinearity, through having two audio features be correllated in some way, could affect the interpretation and relative influence of individual variables on popularity scores.
-Moreover, the use of a linear regression model, while providing good initial insights, could oversimplify the relationships by failing to capture non-linear or simultaneous interaction effects between the variables analyzed.
-Furthermore, the dataset's limited scope, comprising only the top 10 songs per artist and excluding some tracks due to missing audio feature data from Spotify's API, restricts the sample size and may impact the precision and robustness of the parameter estimates.
+Moreover, the use of a linear regression model, while providing initial insights, could oversimplify the relationships by failing to capture non-linear or interaction effects among the variables.
+The dataset's limited scope, comprising only the top 10 songs per artist and excluding some tracks due to missing audio feature data from Spotify's API, restricts the sample size and may impact the precision and robustness of the parameter estimates.
 
-### Next steps
+## Next steps
 
 To address the weaknesses and expand upon the present findings, there are several avenues for future research emerge.
+
 Constructing a more comprehensive dataset with complete artist discographies and a broader array of audio features could better capture the complex relationships between musical characteristics and popularity.
-Exploring non-linear modeling techniques or machine learning algorithms may uncover nuanced patterns and interactions that linear models cannot detect.
-Incorporating alternative data sources or popularity metrics beyond the Spotify platform could provide further insights into the determinants of song appeal and resonance with audiences.
-Repeating the analysis on specific musical genres or using different lists of top artists could uncover genre-specific relationships or other variable influences that may be obscured in the present, more general analysis.
-Finally, validating and cross-referencing the results across multiple data sources would be crucial to assessing their reliability and generalizability to the whole pop music genre.
+Incorporating alternative data sources or popularity metrics beyond the Spotify platform could provide further insights into the determinants of song appeal and resonance with audiences, and could make the results more generalizable to pop music at large.
+That said, few, if any, institutions have the influence or technical capabilities as Spotify to be able to analyze and categorize the vast amounts of audio data present in their vast library.
+Looking into cross-referencing this analysis with other datasets mentioned, like the Million Song Dataset, could be an option for improving the generalizability of the results.
 
-Despite it's shortcomings, my approach affords merit as the first step by a cultural analyst in using the openly available data from Spotify to analyze and quantify what makes popular music popular.
-Through this analysis, I hope to inspire future work in attempting to understand the uniquely human elements of what makes our global culture tick.
+To improve upon the methodological challenges, looking into other linear model designs, such as Bayesian models, could simplify workflows for getting better inferences on updated data.
+This is especially salient given the constant re-calculating of popularity scores by Spotify.
+Exploring non-linear modeling techniques or machine learning algorithms may also uncover nuanced patterns and interactions that simpler linear models cannot detect.
+
+Repeating this analysis on specific musical genres, focusing on particular artists, or using different definitions of success in popular music could also uncover genre-specific relationships or other variable influences that may be still obscured despite in light of me doing this work.
+
+My approach affords merit as the first step by a cultural analyst in using the openly available data from Spotify to analyze and quantify the elements of what could make popular music popular.
+Through this analysis, cultural analysts should be inspired to continue work uncovering the elements that make up popular pop music.
+Artists also should be inspired to create better, more interesting art, rather than be afraid of their work being reduced down to a number; doing statistics is about simplifying, rather than a completely describing the world.
+Although much was learned in this analysis, there is still much to discover about the uniquely human elements of what makes our global culture tick.
 

### 2024-04-25 paper/paper.qmd
final-ish push (just in case I need to backtrack and change something)

@@
 Thank you to Michaela Drouillard for your insights on Spotify and it's metrics.
-Special hat tip goes to Rajan Maghera and my amazing parents, Cristina and Shane for your constant support and encouragement.
+My sincere appreciation goes to Hannah Yu for your valuable last-minute edits.
+Special hat tip goes to Rajan Maghera and my amazing parents, Cristina and Shane for your support and encouragement.
 I could not have done this without you."
@@
 This paper examines the audio features that characterize mainstream music's biggest hits.
-By analyzing data from the discographies of the highest-charting artists on the Billboard Hot 100, the research identifies key attributes associated with popular songs.
+By analyzing data from the discographies of the greatest artists on the Billboard Hot 100, the research identifies key attributes associated with popular songs.
 Multivariate regression analysis reveals that higher levels of danceability, explicit lyrics, and loudness are positively related to popularity, while emotional positivity (valence) exhibits a negative relationship.
-Overall, this work quantifies some attributes underpinning iconic pop successes and can empower professional pop musicians to make more informed creative decisions with their work."
+Overall, this work quantifies some factors that drive iconic pop successes and can empower professional pop musicians to make more informed creative decisions with their work."
 format: pdf
@@
   mutate(`Song Name` = "...") |> 
-  head(5) |>
+  head(3) |>
   arrange(`Artist Name`)
@@
 
-The primarily categorical dataset was then analyzed variable-by-variable.
-Each variable was examined based on its potential impact on popularity scores, measured from 0-100 with 100 being the most popular.
-For further inference, the popularity score was then modeled using multiple linear regression, with the other variables as predictors.
+The dataset was then analyzed variable-by-variable.
+Popularity score was first investigated for potential discernable patterns.
+Then every other variable was examined based on its potential impact on popularity scores.
 
-## Popularity
+## Popularity Score
 
-First, we assess how popularity is distributed across our dataset using a histogram.
+We first calculate summary statistics about the popularity scores in the dataset, shown in @tbl-sumstats-pop.
 
-```{r fig-pop-hist}
+```{r tbl-sumstats-pop}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Summary Statistics of Popularity (0-100)
+#| fig-align: center
+
+summary_stats <- spotify_data |>
+  summarise(
+    Mean = mean(popularity),
+    `Std. Dev` = sd(popularity),
+    Min = min(popularity),
+    `25th percentile` = quantile(popularity, 0.25),
+    Median = median(popularity),
+    `75th percentile` = quantile(popularity, 0.75),
+    Max = max(popularity)
+  )
+
+# Output the table using kable
+kable(summary_stats, digits = 2, format = "html")
+
+```
+
+Then, we assess how popularity is distributed across our dataset using a histogram.
+
+```{r fig-pop-hist, fig.height=3}
 #| message: false
@@
 
-```{r fig-scatter-val}
+We first calculate summary statistics for valence in @tbl-sumstats-val.
+
+```{r tbl-sumstats-val}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Summary Statistics of Valence (measured 0-1)
+#| fig-align: center
+
+summary_stats <- spotify_data |>
+  summarise(
+    Mean = mean(valence),
+    `Std. Dev` = sd(valence),
+    Min = min(valence),
+    `25th percentile` = quantile(valence, 0.25),
+    Median = median(valence),
+    `75th percentile` = quantile(valence, 0.75),
+    Max = max(valence)
+  )
+
+# Output the table using kable
+kable(summary_stats, digits = 2, format = "html")
+
+```
+
+Then, we plot valence scores against popularity scores in a scatterplot in @fig-scatter-val.
+
+```{r fig-scatter-val, fig.height=3}
 #| message: false
@@
 
-When plotting valence values against popularity scores, though, @fig-scatter-val shows no clear linear trend between valence and popularity, although a clumping of high valence scores near 1.00, suggests potential data issues.
-
+When plotting valence values against popularity scores, though, no clear linear trend between valence and popularity makes itself apparent, although a clumping of high valence scores near 1.00, suggests potential data issues.
 ## Danceability
@@
 In their API documentation, Spotify says that "Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
-A value of 0.0 is least danceable and 1.0 is most danceable."
+A value of 0.0 is least danceable and 1.0 is most danceable." @spotifydocs.
+As before, summary statistics are presented in @tbl-sumstats-dance
 
-```{r fig-scatter-danceability}
+```{r tbl-sumstats-dance}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Summary Statistics of danceability (measured 0-1)
+#| fig-align: center
+
+summary_stats <- spotify_data |>
+  summarise(
+    Mean = mean(danceability),
+    `Std. Dev` = sd(danceability),
+    Min = min(danceability),
+    `25th percentile` = quantile(danceability, 0.25),
+    Median = median(danceability),
+    `75th percentile` = quantile(danceability, 0.75),
+    Max = max(danceability)
+  )
+
+# Output the table using kable
+kable(summary_stats, digits = 2, format = "html")
+
+```
+
+```{r fig-scatter-danceability, fig.height=3}
 #| message: false
@@
 In the dataset, 1 indicates the song is major while 0 is minor.
+Due to mode's binary nature, it is best summarized in graphs.
 
@@
 
-```{r fig-mode-density-plot}
+```{r fig-mode-density-plot, fig.height=3}
 #| message: false
@@
 
-Spotify records whether a song contains explicit lyrics through reporting by the music publishers.
+Spotify records whether a song contains explicit lyrics through reporting by the music publishers. Similar to mode, explicitness is best summarized through graphs.
 
@@
 
-```{r fig-explicit-density}
+```{r fig-explicit-density, fig.height=3}
 #| message: false
@@
 
-The average loudness of a song, in decibels (dB) is calculated by averaging the height of the waveforms in a particular song.
+The average loudness of a song, in decibels (dB) is calculated by averaging the height of the waveforms in a particular song.Summary statistics are presented in @tbl-sumstats-loudness.
 In @fig-scatter-loudness we plot loudness against popularity to assess any relationship between the two variables.
 
-```{r fig-scatter-loudness}
+```{r tbl-sumstats-loudness}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Summary Statistics of Loudness (measured in dB)
+#| fig-align: center
+
+summary_stats <- spotify_data |>
+  summarise(
+    Mean = mean(loudness),
+    `Std. Dev` = sd(loudness),
+    Min = min(loudness),
+    `25th percentile` = quantile(loudness, 0.25),
+    Median = median(loudness),
+    `75th percentile` = quantile(loudness, 0.75),
+    Max = max(loudness)
+  )
+
+# Output the table using kable
+kable(summary_stats, digits = 2, format = "html")
+
+```
+
+```{r fig-scatter-loudness, fig.height=3}
 #| message: false
@@
 
+\newpage
 The plot shows no clear linear relationship between loudness and popularity, likely due to the varied nature of popular music and other confounding factors.
-That makes sense, since there could be songs that are "known" for being soft and emotional, while others are loud.
-Given the varied nature of pop music, it is to be expected.
+That makes sense, since there could be songs that are "known" for being soft and emotional, while others are loud and intense.
+Those are two of many descriptions one could give to the multitude of styles and expressions within the dataset.
+Therefore, given the varied nature of pop music, it is to be expected to not observe an overtly linear relationship between solely loudness and popularity score.
 
@@
 However, @fig-dur-scatter reveals no discernible linear relationship between song duration and Spotify popularity score.
+Summary statistics on duration are presented in @tbl-sumstats-duration.
 
-```{r fig-duration-hist}
+```{r tbl-sumstats-duration}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Summary Statistics of Duration (measured in minutes)
+#| fig-align: center
+
+summary_stats <- spotify_data |>
+  mutate(duration_secs = duration_secs/60) |> #convert to minutes
+  summarise(
+    Mean = mean(duration_secs),
+    `Std. Dev` = sd(duration_secs),
+    Min = min(duration_secs),
+    `25th percentile` = quantile(duration_secs, 0.25),
+    Median = median(duration_secs),
+    `75th percentile` = quantile(duration_secs, 0.75),
+    Max = max(duration_secs)
+  )
+
+# Output the table using kable
+kable(summary_stats, digits = 2, format = "html")
+
+```
+
+```{r fig-duration-hist, fig.height=3}
 #| message: false
@@
   geom_histogram(bins = 20, fill = "blue", color = "black") + 
-  labs(title = "Histogram of Song Duration (Mins)",
+  labs(title = "Histogram of Song Duration",
        x = "Duration (mins)",
@@
 
-```{r fig-dur-scatter}
+```{r fig-dur-scatter, fig.height=3}
 #| message: false
@@
 
+\newpage
+
+\newpage 
 # Data Modelling {#sec-model}
 
-While individual variable analyses provide some surface-level insights into the determinants of the popularity of a song, music's complexity necessitates considering variables together for reasonable and informative inference.
+While analyzing individual variables provides surface-level insights into the determinants of the popularity of a song, the complex factors underpinning musics's popularity necessitates accounting for interactions between variables for reasonable and informative inference.
 
 ## Model justification
-Creating a linear regression model is primarily motivated by the fact that it can account for relationships between the variables and the popularity, as well the relationships between the predictor variables themselves.
-Running a linear regression model allows us to isolate effects of a particular variable while also controlling for the effects of the other variables at the same time.
+Creating a linear regression model of this dataset is motivated by the fact that it can account for individual relationships between the audio features and the popularity, as well the relationships between the audio features themselves.
+By running a linear regression model the effects of a particular variable on a Spotify popularity score can be isolated, while also controlling for the effects of the other variables at the same time.
 
-Before conducting the analysis, however, several tests were conducted (shown in [@sec-testing]) to verify the data fit within the linear regression assumptions of linearity, normally distributed errors, and homoscedasticity of residuals.
-This ensured that the data was well suited for analysis using a multiple regression.
+Before conducting the analysis, however, several tests were conducted (shown in @sec-testing) to verify the data fit within the linear regression assumptions of linearity, normally distributed errors, and homoscedasticity of residuals.
+This ensured that the data was well suited for analysis using a multiple regression analysis.
 
@@
 For a particular variable, a low p-value for the regression coefficient would provide evidence against the null hypothesis, indicating that the variable has a meaningful effect on Spotify popularity score.
-On the other hand, if the p-value is high for a given variable, this would suggest that there is not enough evidence to reject the null hypothesis, and there may be no significant linear relationship between that variable and the Spotify popularity score.
+On the other hand, a high p-value for a variable would suggest that there is not enough evidence to reject the null hypothesis, and there may be no significant linear relationship between that variable and the Spotify popularity score.
+
 Ultimately, the results of the analysis will inform whether there is any significant linear relationship, controlling for the other variables,between each the variables and a song's Spotify popularity score.
 
-The aim of computing a regression model is to estimate a line with the coefficients $\beta_0$ to $\beta_6$ so that the difference (error) between the predicted line and the data points is minimized.
-By doing this, we get an equation for a line that best fits the data, allowing for the estimation of a Spotify score for a given set of audio features.
+The aim of computing a regression model is to estimate the equation for a line with the coefficients $\beta_0$ to $\beta_6$ so that the difference (error) between the predicted line and the data points is minimized.
+By doing this, we get an equation for a line that best fits the data, allowing for the estimation of a Spotify popularity score for a given set of audio features.
+With a multiple regression, we use multiple different predictors (or independent variables) to predict a single outcome variable.
 
-The equation for our linear model can be written as follows:
+The equation for our linear model can be written as
 
@@
 
-- The variable $V$ denotes the valence (emotional positivity) of the song, measured from 0 to 1, where 1 is the most positive.
+- $V$ denotes the valence (emotional positivity) of the song, measured from 0 to 1, where 1 is the most positive.
 The coefficient $\beta_1$ quantifies how a one-unit increase in valence (e.g., from 0.5 to 1.0) affects the popularity score $S$.
 
-- The variable $D$ represents the danceability of the song, also scaled from 0 to 1, with higher values indicating more danceability.
+- $D$ represents the danceability of the song, also measured from 0 to 1, with higher values indicating more danceability.
 The coefficient $\beta_2$ captures the change in $S$ for a one-unit increase in danceability.
 
-- The binary variable $M$ indicates the musical mode (0 for minor, 1 for major), with $\beta_3$ reflecting the difference in popularity between major and minor keys.
+- $M$, a binary variable indicates the musical mode (0 for minor, 1 for major), with $\beta_3$ reflecting the difference in popularity between major and minor keys.
 
-- The binary variable $E$ indicates explicit lyrics (0 for non-explicit, 1 for explicit), where $\beta_4$ estimates the effect of explicit content on popularity.
+- $E$, another binary variable, indicates explicit lyrics (0 for non-explicit, 1 for explicit), where $\beta_4$ estimates the effect of explicit content on popularity.
 
-- The variable $L$ denotes the loudness of the song in decibels, with $\beta_5$ quantifying how a one-decibel increase in loudness impacts the popularity score $S$.
+- $L$ denotes the loudness of the song in decibels, with $\beta_5$ quantifying how a one-decibel increase in loudness impacts the popularity score $S$.
 
-- Finally, $T$ (for time), represents the duration of the song in seconds, and $\beta_6$ estimates the effect of an additional second on the popularity score.
+- $T$ (for time), represents the duration of the song in seconds, and $\beta_6$ estimates the effect of an additional second on the popularity score.
 
@@
   Term = rownames(model_summary$coefficients),
-  Estimate = model_summary$coefficients[, "Estimate"],
+  `Coefficient Estimate` = model_summary$coefficients[, "Estimate"],
   `Std. Error` = model_summary$coefficients[, "Std. Error"],
@@
 
-```{r effect-plot-valence, fig.height=3}
+```{r fig-effect-plot-valence, fig.height=3}
 #| message: false
@@
 #| warning: false
-#| fig-cap: Effect Plot of Valence on Popularity
+#| fig-cap: "Effect Plot of Valence on Popularity"
 #| fig-align: center
@@
 
-Valence, which measures the emotional positivity of a song on a scale of 0 to 1, has a negative coefficient (-10.96).
-This suggests that, holding other variables constant, an increase in valence by one unit is associated with a decrease in the predicted Spotify popularity score by approximately 11 points.
+Valence (shown in @fig-effect-plot-valence), which measures the emotional positivity of a song on a scale of 0 to 1, has a negative coefficient (-10.96).
+This suggests that when comparing two songs that are the same in danceability, mode, explicit lyrics, loudness, and duration, the song with a valence (emotional positivity) score that is 1 unit higher is predicted to have a Spotify popularity score that is 11 points lower, on average.
 
-```{r effect-plot-danceability, fig.height=3}
+```{r fig-effect-plot-danceability, fig.height=3}
 #| message: false
@@
 #| warning: false
-#| fig-caption: Effect Plot of Danceability on Popularity
+#| fig-cap: "Effect Plot of Danceability on Popularity"
 #| fig-align: center
@@
 
-Danceability, also scaled from 0 to 1, has a positive coefficient (18.24).
-This indicates that, controlling for other factors, a one-unit increase in danceability is associated with an increase in the predicted popularity score by about 18 points.
+Danceability (shown in @fig-effect-plot-danceability), also scaled from 0 to 1, has a positive coefficient (18.24).
+This means that when comparing two songs that have the same valence, mode, explicit lyrics, loudness, and duration, the song with a danceability score that is 1 unit higher is predicted to have a Spotify popularity score that is 18 points higher, on average.
 
-```{r effect-plot-mode, fig.height=3}
+```{r fig-effect-plot-mode, fig.height=3}
 #| message: false
@@
 #| warning: false
+#| fig-cap: "Effect Plot of Musical Mode on Popularity"
 #| fig-align: center
@@
 
-Explicit is another binary variable (possibly indicating the presence of explicit lyrics), with a positive coefficient (10.10).
-This implies that, holding other variables constant, songs with explicit lyrics are predicted to have a popularity score that is approximately 10 points higher than non-explicit songs.
+Mode is a binary variable (0 being minor, 1 being major), with a negative coefficient (-0.85), but has a p-value of 0.37, which indicates no statistical significance.
+Since the p-value is too high for our alternative hypothesis to be satisfied, we revert to our null hypothesis.
+We conclude that an average song's musical mode does not have a significant effect on the song's Spotify popularity score.
+I visualize this in @fig-effect-plot-mode.
 
-```{r effect-plot-explicit, fig.height=3}
+```{r fig-effect-plot-explicit, fig.height=3}
 #| message: false
@@
 #| warning: false
+#| fig-cap: "Effect Plot of Explicit Lyrics on Popularity"
 #| fig-align: center
@@
 
-Loudness, measured in decibels, has a positive coefficient (0.82), indicating that an increase in loudness by one decibel is associated with an increase in the predicted popularity score by 0.82 points, controlling for other factors.
+Explicit (shown in @fig-effect-plot-explicit) is another binary variable (indicating the presence of explicit lyrics), with a positive coefficient (10.10).
+This implies that when comparing two songs that are the same in other variables, the song that contains explicit lyrics is predicted to have a Spotify popularity score that is 10 points higher, on average, compared to a song without explicit lyrics.
 
-```{r effect-plot-loudness, fig.height=3}
+```{r fig-effect-plot-loudness, fig.height=3}
 #| message: false
@@
 #| warning: false
+#| fig-cap: "Effect Plot of Loudness on Popularity"
 #| fig-align: center
@@
 
-Duration has a positive coefficient (0.01) with a p-value of 0.10, which is marginally significant.
-This suggests that, holding other variables constant, an increase in song duration by one second is associated with a slight increase in the predicted popularity score by 0.01 points.
+Loudness (shown in @fig-effect-plot-loudness), measured in decibels, has a positive coefficient (0.82).
+We can say that when comparing two songs that have the same in other variables, the song with an average loudness level that is 1 decibel higher is predicted to have a Spotify popularity score that is 0.82 points higher, on average.
 
-```{r effect-plot-duration, fig.height=3}
+```{r fig-effect-plot-duration, fig.height=3}
 #| message: false
@@
 #| warning: false
+#| fig-cap: "Effect Plot of Song Duration on Popularity"
 #| fig-align: center
@@
 
-Taken together, the regression results indicate that danceability, explicit lyrics, loudness, and, to a lesser extent, song duration are associated with higher Spotify popularity scores, while valence (emotional positivity) has a negative association.
-The musical mode does not appear to have a significant effect on popularity after accounting for other predictors.
+Duration (@fig-effect-plot-duration) has a positive coefficient (0.01) with a p-value of 0.10, which indicates the coefficient has low statistical significance.
+We can say then, with low confidence, that when comparing two songs that have the same levels of valence, danceability, mode, explicit lyrics, and duration, the song with a duration 1 second longer is predicted to have a Spotify popularity score that is 0.01 points higher, on average.
+
+Taken together, the regression results indicate that higher Spotify popularity scores are associated with danceability, explicit lyrics, loudness, and to a lesser extent, song duration, while valence (emotional positivity) has a negative association.
+The musical mode does not appear to have a significant effect on popularity when other predictors are accounted for.
 
@@
 These findings offer valuable insights into the extensive efforts undertaken by music technology companies like Spotify to quantify and understand the multifaceted dynamics of music culture.
-The analysis underscores the potential impact and utility of quantitative, data-driven methods in comprehending and potentially shaping cultural phenomena like popular music.
+The analysis underscores the potential impact and utility of quantitative, data-driven methods in comprehending and shaping cultural phenomena like popular music.
 
@@
 
-There are weaknesses and limitations in my research that warrant consideration.
+There are weaknesses and limitations in this research that warrant consideration.
 The age of the data used, with the Billboard Greatest of All Time list concluding in 2015, may not fully capture contemporary trends and shifts in music culture, especially given the rapid rise of social media platforms like TikTok that drive rapid virality and micro-trends in popular music consumption.
 
-Additionally, the crowdsourced nature of certain audio features like valence and danceability, initially derived from subjective assessments by college interns, introduces an element of human subjectivity and potential bias, which could undermine the validity and generalizability of the relationships observed between these features and popularity scores.
+Additionally, the crowdsourced nature of certain audio features like valence and danceability, initially derived from subjective assessments by college interns @drouillard2024, introduces a clear element of human subjectivity and potential bias in the collection of data, which could undermine the validity and generalizability of the relationships observed between these features and popularity scores.
 The reliance solely on Spotify popularity scores as the dependent variable could bias the results toward newer music versus older, more established releases, although the recency factor incorporated into Spotify's scoring algorithm could potentially mitigate this concern for tracks that have maintained enduring popularity over time.
-Since the mechanics of calculating the Spotify popularity score are proprietary, there is
+In particular, given the little amount of open and reliable information about how the Spotify popularity score and other generated metrics are calculated, it is worth questioning whether if those scores are associated with popularity scores by design.
 
@@
 Most obviously, potential multicollinearity among the predictor variables (audio features) could affect the interpretation and relative influence of individual variables on popularity scores.
-In particular, given the little amount of reliable information about how a Spotify popularity score is calculated, it is worth questioning whether if certain algorithmically generated scores such as danceability are associated with popularity scores by design.
 
@@
 
-# Appendix {-}
+# Appendix {#sec-appendix}
 
@@
 ```
-
-```{r boilerplate-r-chunk}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: CAPTION
-#| fig-align: ALIGNMENT
-
-  theme(plot.title.position = "plot",
-        plot.title = element_text(hjust = 0.5))
-
-```

### 2024-04-25 scripts/00-simulate_data.R

@@ -32,12 +32,9 @@ simulated_data <- tibble(
   popularity = sample(0:100, num_songs, replace = TRUE),
   song = stri_rand_strings(num_songs, 10), # Random strings of length 10 for song titles
   artist = stri_rand_strings(num_songs, 7), # Random strings of length 7 for artist names
-  energy = runif(num_songs, 0, 1), # Uniform distribution between 0 and 1
   valence = runif(num_songs, 0, 1),
   danceability = runif(num_songs, 0, 1),
-  liveness = runif(num_songs, 0, 1),
   explicit = sample(0:1, num_songs, replace = TRUE), # Binary variable, assuming 0 is not explicit, 1 is explicit
-  instrumentalness = runif(num_songs, 0, 1),
   duration_ms = sample(150000:300000, num_songs, replace = TRUE), # Duration in milliseconds, example range
   mode = sample(0:1, num_songs, replace = TRUE) # Binary variable, assuming 0 is minor, 1 is major
 )
@@ -74,7 +71,7 @@ print(c(duplicated_songs, duplicated_artists))
 
 # 5. Range Validation Test
 range_checks <- sapply(
-  simulated_data[c("energy", "valence", "danceability", "liveness", "instrumentalness")],
+  simulated_data[c("valence", "danceability")],
   function(x) all(x >= 0 & x <= 1)
 )
 print(range_checks)

### 2024-04-25 scripts/03-test_data.R

@@ -48,5 +48,13 @@ test_that("There are no missing values in key columns", {
   }
 })
 
+test_that("Loudness values are within realistic ranges", {
+  expect_true(all(cleaned_ds$loudness >= -60 & cleaned_ds$loudness <= 0))
+})
+
+test_that("Duration of songs is within plausible limits", {
+  expect_true(all(cleaned_ds$duration_secs >= 30 & cleaned_ds$duration_secs <= 1500))
+})
+
 # You can run the tests with the following line
 test_dir("tests/")

### 2024-04-25 paper/paper.qmd
edited paper (for the last time, hopefully)

@@
 My sincere appreciation goes to Hannah Yu for your valuable last-minute edits.
-Special hat tip goes to Rajan Maghera and my amazing parents, Cristina and Shane for your support and encouragement.
-I could not have done this without you."
+Special hat tips go to Rajan Maghera and my amazing parents, Cristina and Shane for your support and encouragement.
+I could not have done this without you all of your support."
 date: April 24 2024
@@
 
-Taken together, the regression results indicate that higher Spotify popularity scores are associated with danceability, explicit lyrics, loudness, and to a lesser extent, song duration, while valence (emotional positivity) has a negative association.
-The musical mode does not appear to have a significant effect on popularity when other predictors are accounted for.
+The regression results indicate that higher Spotify popularity scores are associated with danceability, explicit lyrics, loudness, and to a lesser extent, song duration, while valence (emotional positivity) has a negative association.
+On the other hand, musical mode does not appear to have a significant effect on popularity when other predictors are accounted for.
 
@@
 They could also choose to diverge from these elements in order to explore new genres or music styles that may not be popular now, but could be in the future.
-Knowing the current trends now could inspire artists to make better art that effectively and innovatively builds on what has come before.
+Knowing the current trends now could inspire artists to make better art that effectively innovates on what has come before.
 
@@
 Additionally, the crowdsourced nature of certain audio features like valence and danceability, initially derived from subjective assessments by college interns @drouillard2024, introduces a clear element of human subjectivity and potential bias in the collection of data, which could undermine the validity and generalizability of the relationships observed between these features and popularity scores.
-The reliance solely on Spotify popularity scores as the dependent variable could bias the results toward newer music versus older, more established releases, although the recency factor incorporated into Spotify's scoring algorithm could potentially mitigate this concern for tracks that have maintained enduring popularity over time.
-In particular, given the little amount of open and reliable information about how the Spotify popularity score and other generated metrics are calculated, it is worth questioning whether if those scores are associated with popularity scores by design.
+The reliance solely on Spotify popularity scores as the dependent variable could bias the results toward newer music versus older, more established releases; Spotify is a digital-first company, meaning that their records do not include performance of songs that were released before their audio feature API went live in the early 2010s.
+That said, the recency factor incorporated into Spotify's scoring algorithm for popularity could potentially mitigate this concern for tracks that have maintained enduring popularity over time.
+Given the little amount of open and reliable information about how the Spotify popularity score and other generated metrics are calculated, there is much to question about the creation of these metrics, even considering whether these metrics were meant to be associated by design.
 
-There could also be potential methodological issues stemming from the complex and interrelated nature of the data itself.
+There are possible methodological issues in this design, primarily stemming from the complex and interrelated nature of the data itself.
 Most obviously, potential multicollinearity among the predictor variables (audio features) could affect the interpretation and relative influence of individual variables on popularity scores.
-
-Moreover, the use of a linear regression model, while providing initial insights, could oversimplify the relationships by failing to capture non-linear or interaction effects among the variables.
-The dataset's limited scope, comprising only the top 10 songs per artist and excluding some tracks due to missing audio feature data from Spotify's API, restricts the sample size and may impact the precision and robustness of the parameter estimates.
+The use of a linear regression model, while providing initial insights, could have oversimplified relationships by failing to capture non-linear or interaction effects among the predictor variables.
+Given the dataset's limited scope, comprising only the top 10 songs per artist and excluding some tracks due to missing audio feature data from Spotify's API, means that the precision and robustness of the parameter estimates could likely be inaccurate and not generalizable.
 
@@
 
-To address the weaknesses and expand upon the present findings, there are several avenues for future research emerge.
+To address the weaknesses and expand upon the present findings, there are many, many emerging avenues for future research in this area.
 
@@
 
-Repeating this analysis on specific musical genres, focusing on particular artists, or using different definitions of success in popular music could also uncover genre-specific relationships or other variable influences that may be still obscured despite in light of me doing this work.
+Finally, Repeating this analysis on specific musical genres, focusing on particular artists, or using different definitions of success in popular music could also uncover genre-specific relationships or other variable influences that may be still obscured despite in light of me doing this work.
 
@@
 Artists also should be inspired to create better, more interesting art, rather than be afraid of their work being reduced down to a number; doing statistics is about simplifying, rather than a completely describing the world.
-Although much was learned in this analysis, there is still much to discover about the uniquely human elements of what makes our global culture tick.
+Although much was learned in this analysis, there is still much to discover about the uniquely human elements of what makes our globally connected culture tick.
 
